### 2025-01-30 src/paper/paper.tex
updated paper

\input{preamble}

\begin{document}

% Title
\maketitle

% Abstract
\section{Abstract}

\textbf{Background}\\
Schizophrenia (SZ) is a chronic mental disorder that presents challenges in accurate diagnosis due to its complex nature. Advances in machine learning (ML) and deep learning (DL) have shown promise in automating SZ detection using structural MRI, but variability in preprocessing methods remains a key challenge.
\textbf{Aim}: This study aims to evaluate the impact of various preprocessing and data augmentation techniques on ML-based classification of schizophrenia from MRI data.

\textbf{Methods}\\
We utilized the COBRE dataset, which includes structural MRI data from 62 individuals, divided into 30 schizophrenia patients and 32 healthy controls. A support vector classifier (SVC) was employed to classify the data. Various preprocessing techniques—normalization, brain extraction, and cropping—and data augmentation methods—translation, rotation, and Gaussian noise addition—were applied to the MRI images. Model performance was assessed using accuracy, recall, precision, and F1-score.

\textbf{Results}\\
Normalization and brain extraction improved classification accuracy, with cropping achieving the highest accuracy (0.57) and F1-score (0.69) among the preprocessing techniques. Combining preprocessing methods did not always result in performance improvements and sometimes led to a decrease in accuracy. Data augmentation significantly enhanced model performance, with translation and rotation leading to the highest accuracy (0.79) and F1-score (0.80). However, Gaussian noise had a detrimental effect, reducing accuracy and F1-score.

\textbf{Conclusion}\\
Our findings indicate that careful selection of preprocessing and augmentation techniques is critical for optimizing model performance in schizophrenia detection. While certain preprocessing steps like normalization and cropping are beneficial, combining them may not always yield additive improvements. Augmentation techniques like translation and rotation proved effective in boosting model performance, while Gaussian noise should be avoided. These results contribute to the development of more standardized preprocessing pipelines for medical imaging, which could enhance the reliability and reproducibility of machine learning models in clinical applications.

\textbf{Keywords: schizophrenia, sMRI, CNN, SVC}

% Introduction
\section{Introduction}

Schizophrenia (SZ) is a chronic mental disorder affecting 1 in 300 people worldwide, linked to structural brain abnormalities observable via MRI. Machine learning (ML) and deep learning (DL) have advanced SZ diagnosis, yet variability in preprocessing methods—such as noise removal and intensity normalization—remains a major challenge~\cite{Benli2023, Zhang2022}.

Despite progress in CNN-based SZ detection, studies often lack standardized preprocessing pipelines, impacting model reliability and reproducibility. Benchmark datasets~\cite{Oh2020, Verma2023} report high classification accuracies, but inconsistent preprocessing limits comparability.

This project systematically evaluates preprocessing and augmentation techniques (e.g., smoothing, translation) to establish a robust pipeline for SZ detection from MRI. Standardizing these practices could improve the reproducibility and clinical applicability of ML-based diagnostics.

% Methods
\section{Methods}

\subsection{Dataset and Exploratory Data Analysis}

The raw data were obtained from the \href{http://schizconnect.org}{SchizConnect} database (accessed January 1, 2025), which houses structural and functional MRI data. The database provides filtering options to enable users to select data that meets specific criteria including MRI field strength and clinical diagnosis (e.g. schizophrenia, bipolar disorder).

The data obtained originates from the Center of Biomedical Research Excellence (COBRE) dataset in NifTI format. It includes data from 62 individuals, 30 with schizophrenia and 32 healthy controls.

The data from SchizConnect was filtered based on MRI field strength and clinical diagnosis. We retrieved only images captured with 3T MRI scans. This was done to ensure uniformity of the data obtained.

Clinical data was made available in form of CSV files. The demographic features of the dataset used in this project are presented in Table~\ref{tab:cobre_clinical_demographic}.
\begin{center}
	\begin{table}
        \centering
        \caption{\label{tab:cobre_clinical_demographic}The subjects were selected to give a balanced representation of age and gender distributions in the dataset, prioritizing age distribution, which has been shown to affect model performance (ref).}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}lcccccc@{\extracolsep\fill}}
            \toprule
            & \multicolumn{3}{c}{COBRE}
            \\\cmidrule{2-4}
            & \textbf{Individuals with SZ (n=30)} & \textbf{Healthy controls (n=32)} & \textbf{Total (62)} \\
            \midrule
            Minimum age             & 19  & 18  & 18 \\
            Maximum age 		    & 66  & 65 & 66  \\
            Average age             & 39.6 & 41 & 40.3	\\
            Gender (Male/ Female)   & 19/11 & 17/15 & 36/26	\\
            \bottomrule
        \end{tabular*}
    \end{table}    
\end{center}

%
\subsection{Image Preprocessing}

% General thoughts
The major idea behind preprocessing was to establish a standardized pipeline that ensures consistency in the quality and content of the MRI data. This pipeline addressed common issues such as noise, intensity variations, and artifacts, which can affect the performance of machine learning models.

We experimented with different combinations of preprocessing and data augmentation techniques to determine which pipeline yields the best performance of the ML/DL models regarding recall rate, precision, and F1 score.

In order to find the best combination of hyperparameters for each preprocessing method, we performed a grid search and measured the quality metrics of the processed image using a custom test function.

Table~\ref{tab:preprocessing_pipeline} presents the preprocessing techniques used and their respective hyperparameters and tools used. Figure~\ref{fig:test_pipeline} gives a visual representation of the conducted tests.
%
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{./figs/empty.pdf} % test_pipeline.png
    \caption{Tests pipeline for MRI data.}\label{fig:test_pipeline}
\end{figure}

\subsubsection{Resampling}

The first step of the pipeline is resampling the image to a standard resolution to ensure uniform input for the model and consistent noise and signal profiles.

\subsubsection{intensity normalization}

The second step of the pipeline is normalization of voxel intensities to a consistent range to maintain the relative scale of the data and produce values in a fixed range to satisfy models expectations. The adopted method scales the data to a range of [0, 1].

Normalization also improves the results of brain extraction by helping the algorithm better distinguish between brain and non-brain regions.

Intensity distributions were inspected to ensure normalization has not distorted the data.

\subsubsection{Brain extraction}

The third step of the pipeline is a brain extraction. The purpose of this stage is to
\begin{itemize}
    \item remove non-brain regions in order to minimizes noise, making it easier for the network to focus on relevant features;
    \item subsequently enhance the network's performance, especially in a schizophrenia detection task where small differences in brain anatomy are crucial;
    \item ensure that the input images are more consistent in terms of content, which can help the model generalize better.
\end{itemize}

The generated brain mask was refined by closing gaps and removing small regions. The mask was also used as a leverage for the cropping step.

\subsubsection{Cropping}

The fourth step of the pipeline is cropping the brain region to a consistent size across scans.

\subsubsection{Smoothing}

The fifth step of the pipeline is smoothing the brain region to reduce noise and variability introduced by acquisition artifacts, enhancing the overall quality of the images. This was done to ensure that the downstream tasks are more robust and the model training is more effective. We used a Gaussian filter to reduce high-frequency noise while preserving edges to some extent. Since the neural network performed a classification task, moderate smoothing was carried out to focus on broader patterns.

The sixth step of the pipeline is normalization of the smoothed image to address potential intensity changes and ensure that all images are on a comparable scale.

\subsubsection{Validation}

The impact of preprocessing on the data's integrity and quality was assessed using a combination of a visual inspection and quantitative metrics. The visual inspection involved plotting a few slices of the MRI before and after each preprocessing step. Validation was performed to ensure that preprocessing steps did not introduce artifacts or distort anatomical structures, and to ensure consistency of the processed images.

Quantitative metrics included Signal-to-Noise Ratio (SNR), Contrast-to-Noise Ratio (CNR), Peak Signal-to-Noise Ratio (PSNR), and 
Structural Similarity Index (SSIM).

Relative PSNR for one image. Using a hypothetical "ideal" signal (the maximum intensity value as a baseline), a relative PSNR for one image was calculated. This relative PSNR was used as a way to assess the "quality" of a single image compared to a theoretical perfect image of uniform intensity.

A binary mask was computed using Otsu's thresholding to explicitly select a background region for noise estimation.

The metrics were computed for the entire 3D volume to assess the overall quality or similarity of the entire MRI dataset, capturing global differences.
\begin{center}
    \begin{table}
        \centering
        \caption{\label{tab:preprocessing_pipeline}Preprocessing techniques used in the study.}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}
            \toprule
            \textbf{Preprocessing Step} & \textbf{Hyperparameters} & \textbf{Tools} \\
            \midrule
            Resampling & Voxel size: \texttt{2x2x2} & Nibabel library \\
            Intensity Normalization & Min-max scaling: [0, 1] & NumPy \\
            Brain Extraction & Modality: T1, U-net & ANTsPyNet \\
            Cropping & Size: based on the largest bounding box among all images & scikit-image, NumPy \\
            Smoothing & Gaussian filter: sigma=0.5 & SciPy \\
            \bottomrule
            \end{tabular*}
    \end{table}
\end{center}
%
\subsection{Data Augmentation}

According to typical variations in brain MRI scans, the following transformations with corresponding parameters were applied to the images:
\begin{itemize}
    %\item[\textbf{Option 1}]  \textbf{Shearing.} A shearing angle is selected randomly within the range $\left[-15^{\circ}, 15^{\circ}\right]$ and A shearing operation is applied within the x and y-axis. This operation is repeated 10 times with 10 different shearing angles. By this method, 10 images are generated from an original image;
    \item[\textbf{Option 1}] \textbf{Translation}. The operation is applied on the x and y-axis and repeated 10 times by using different values sampled randomly within the range $\left[-10, 10\right]$. By this method, 10 images are generated from an original image.
    \item[\textbf{Option 2}] \textbf{Rotation.} Rotation angles are selected randomly within the range  $\left[-10^{\circ}, 10^{\circ}\right]$ and clockwise rotations are applied using 10 different angle values. 10 images are generated from an original image by this method.
    \item[\textbf{Option 3}] \textbf{Gaussian noise addition.} Noise addition is applied by \mbox{$I=I+R*I*std+mean$}, where $I$ refers to the image, $R$ refers to the Noise. The noise to be added is generated using a Gaussian distribution. The variance of the distribution is computed from the input images while the mean value is set to zero. In this study, three different fixed values (0.1, 0.2, and 0.3) are used and 3 images are generated from an input image.
    %\item[\textbf{Option 4}] \textbf{Gaussian noise addition followed by rotation.} 30 images are generated from an input image by this method.
    %\item[\textbf{Option 5}] \textbf{Clockwise rotation followed by translation.} Rotation angles are selected randomly within the range $\left[-25^{\circ}, 25^{\circ}\right]$ and translation is applied on the x and y-axis by using different values sampled randomly within the range $\left[-15^{\circ}, 15^{\circ}\right]$. This successive operation is applied 10 times, and 10 images are generated from an input image by this method.
    %\item[\textbf{Option 6}] \textbf{Translation followed by shearing.} The translation step is applied on the x and y-axis with different values sampled randomly within the range $\left[-15, 15\right]$. In the shearing step, an angle is selected randomly within the range $\left[-15^{\circ}, 15^{\circ}\right]$, and shearing is applied within the x and y-axis. The subsequent operations are applied 10 times, and 10 images are generated from an input image by this method.
    %\item[\textbf{Option 7}] \textbf{Translation, shearing, and rotation.} Operations are applied subsequently. Implementation of the translation and shearing operations are performed as explained in the previous method. In the rotation step, a random angle is selected within the range $\left[-25^{\circ}, 25^{\circ}\right]$ clockwise rotation is applied. The successive operations are applied 10 times, and 10 images are generated from an input image by this method.
\end{itemize}

The augmented samples were plotted to ensure the transformations don't distort anatomical features.

%
\subsection{Model architectures}

Convolutional Neural Networks (CNNs) are commonly used in the literature for this task. However, a few papers explored the use of support vector machines for image classification. To gain a full understanding of both approaches, we decided on the combination of a pretrained deep neural network for feature extraction and a support vector classifier for binary classification. The ReNet-18 architecture was selected for its balance between computational load and accuracy. Cross validation was performed with the support vector classifier (SVC) to choose the best hyperparameters, and the hyperparameters that yielded the best results were the radial basis function (RBF) kernel, a gamma value of 0.0001 and a C parameter value of 100. Each 2D slice of the images were fed into the ResNet-18 model as NumPy arrays of dimension $224x224$. This input size corresponds to the size of the images in the ImageNet dataset that was used to train ResNet-18 and was selected for compatibility with the pretrained weights of the ResNet-18 model. Figure~\ref{fig:training_architecture} gives a representation of the training architecture and flow.
%
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{./figs/model_architecture.png} % training_architecture.png
    \caption{Training architecture and flow.}\label{fig:training_architecture}
\end{figure}

%
\subsection{Training and validation}

The dataset was split into a training set (80\%) and a validation set (20\%).

Accuracy, recall, precision, and F1-score were calculated to asses the performance of the SVC.

% Results
\section{Results}

\subsection{Image Preprocessing}

The results presented in Table~\ref{tab:model_metrics_preprocessing} demonstrate the impact of various preprocessing methods on classification performance. The baseline model using raw images (0) achieved an accuracy of 0.51, with an F1-score of 0.63. Applying resampling (1) alone did not alter these results, suggesting that resampling alone does not introduce substantial benefits.

Normalization (2) improved accuracy to 0.56 and F1-score to 0.67, indicating that standardizing intensity values enhances model performance. Brain extraction (3) yielded a slightly lower accuracy (0.54) but a notable increase in recall (0.78), which suggests that removing non-brain tissues helps capture more positive instances but may introduce a trade-off in precision.

Cropping (4) led to the highest accuracy (0.57) and an F1-score of 0.69, highlighting its effectiveness in improving signal-to-noise ratio. Smoothing (5) showed a moderate impact, improving results compared to the raw data but not outperforming other methods.

When combining preprocessing techniques, the pattern is less straightforward. The combination of resampling, normalization, and brain extraction (1+2+3) resulted in the highest recall (0.80) while maintaining a strong F1-score (0.69). However, adding cropping (1+2+3+4) caused a drastic performance drop, with accuracy declining to 0.44 and F1-score to 0.54. The full pipeline (1+2+3+4+5) slightly improved over this but still underperformed compared to individual preprocessing steps.
\begin{center}
	\begin{table}[t]
		\centering
		\caption{\label{tab:model_metrics_preprocessing}Desct}
		\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
			\toprule
		     \textbf{Preprocessing method} &\textbf{Accuracy} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
			\midrule
			Raw (0) & 0.51 & 0.67 & 0.60 & 0.63 \\
			Resampling (1) & 0.51 & 0.67 & 0.60 & 0.63 \\
			Normalization (2) & 0.56 & 0.69 & 0.64 & 0.67 \\
			Brain extraction (3) & 0.54 & 0.78 & 0.60 & 0.68 \\
			Cropping (4) & 0.57 & 0.76 & 0.63 & 0.69 \\
			Smoothing (5) & 0.55 & 0.71 & 0.62 & 0.66 \\
            1+2 & 0.57 & 0.69 & 0.64 & 0.67 \\
            1+2+3 & 0.55 & 0.80 & 0.61 & 0.69 \\
            1+2+3+4 & 0.44 & 0.53 & 0.56 & 0.54 \\
            1+2+3+4+5 & 0.48 & 0.60 & 0.58 & 0.59 \\
			\bottomrule
		\end{tabular*}
	\end{table}
\end{center}

\subsection{Data Augmentation}

The results presented in Table~\ref{tab:model_meetrics_augmentation} highlight the impact of various data augmentation methods on model performance. The baseline model using raw data (0) achieved an accuracy of 0.51 and an F1-score of 0.63. In contrast, applying translation (1) significantly improved both accuracy (0.75) and F1-score (0.79), with a notable increase in recall (0.84), indicating that translation is effective at boosting model performance and sensitivity.

Rotation (2) further improved results, with the highest accuracy (0.79) and F1-score (0.80), alongside strong recall (0.81) and precision (0.78), demonstrating its ability to enhance both model robustness and overall performance.

Gaussian noise (3), however, led to a substantial decrease in performance, with an accuracy of only 0.33 and a low F1-score of 0.40. This suggests that adding noise may negatively affect the model, potentially obscuring relevant patterns in the data.

\begin{center}
	\begin{table}[t]
		\centering
		\caption{\label{tab:model_meetrics_augmentation}Desct}
		\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
			\toprule
			\textbf{Augmentation method} &\textbf{Accuracy} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
			\midrule
			Raw (0) & 0.51 & 0.67 & 0.60 & 0.63 \\
			Translation (1) & 0.75 & 0.84 & 0.74 & 0.79 \\
			Rotation (2) & 0.79 & 0.81 & 0.78 & 0.80 \\
			Gaussian Noise (3) & 0.33 & 0.32 & 0.53 & 0.40 \\
			\bottomrule
		\end{tabular*}
	\end{table}
\end{center}

% Discussion
\section{Discussion}

This study investigated the effects of various preprocessing and data augmentation techniques on the classification performance of schizophrenia from structural MRI data. The results demonstrate that certain preprocessing steps, such as normalization, brain extraction, and cropping, can improve model performance when applied individually. However, combining these steps did not yield consistent improvements, suggesting that over-processing may distort the data and hinder the model's ability to generalize effectively.

Among the individual preprocessing techniques, normalization resulted in the most substantial performance improvement, enhancing accuracy by 9.8\%. This can be attributed to the equalization of feature contributions, which allowed the model to make more accurate predictions. Interestingly, this outcome contradicts the conventional understanding that normalizing medical images may flatten clinically significant differences in intensity. The effectiveness of this technique in this study highlights the importance of context when applying preprocessing methods to medical imaging.

Brain extraction, or skull stripping, while beneficial in other studies, had a moderate effect here, improving recall but slightly lowering accuracy. This suggests that removing non-brain tissues enhances the model's sensitivity to schizophrenia cases but may reduce precision, potentially due to the loss of relevant contextual information in the brain image. Cropping, on the other hand, demonstrated a clear advantage by improving accuracy and F1-score. By focusing on the brain region and reducing irrelevant background data, cropping effectively improved the signal-to-noise ratio, allowing the model to focus on clinically relevant features.

However, combining preprocessing steps often led to diminished performance, especially when all techniques were applied together. The full preprocessing pipeline, including resampling, normalization, brain extraction, cropping, and smoothing, resulted in a performance drop, particularly in terms of accuracy and F1-score. This suggests that certain preprocessing steps, when combined, may interfere with each other, creating distortions that prevent the model from learning the most pertinent features of the data.

Regarding data augmentation, both translation and rotation significantly enhanced model performance, with translation improving recall and rotation achieving the highest accuracy and F1-score. These results align with the literature, where such transformations are often used to increase the diversity of the training data, helping the model generalize better to unseen data. Gaussian noise, however, severely degraded performance, suggesting that it may obscure important features in the MRI images, leading to poorer model predictions.

While preprocessing techniques like skull stripping, cropping, and smoothing have been shown to improve performance in various medical imaging tasks, the results in this study suggest that their effectiveness depends on the context and the combination of other preprocessing steps. Furthermore, the adverse effect of Gaussian noise indicates the need for careful consideration of the types of augmentation applied, as certain methods may introduce detrimental effects.

In contrast to some previous studies that have used more complex augmentation techniques, such as elastic deformations, this study focused on simpler methods, which proved to be more effective. Elastic deformations can sometimes introduce unrealistic and unnatural variations, leading to damage and noise that harm model performance~\cite{Mok2018}. The methods investigated in this study, particularly translation and rotation, were found to be effective in maintaining the integrity of the original images while enhancing model robustness.

Overall, this study highlights the importance of standardizing preprocessing and augmentation techniques in medical imaging, particularly for tasks like schizophrenia detection. While certain methods demonstrated clear benefits, the combination of multiple preprocessing steps should be approached with caution, as it may lead to diminishing returns or reduced model performance. Further research is needed to explore the optimal combination of preprocessing techniques and to investigate the impact of more advanced augmentation methods on model accuracy and generalizability.

%augmentation with the combination of shearing and salt-and-pepper noise addition is the least efficient approach for augmentation in improving classification performance

% Discuss When Brain Extraction May Not Be Necessary

% Discuss Metrics for Each 2D Slice (Slice-wise Approach):

% Conclusion
\section{Conclusion}

In this study, we investigated the impact of various preprocessing and data augmentation techniques on the classification performance of schizophrenia from structural MRI images. Our results demonstrate that individual preprocessing methods, particularly normalization, brain extraction, and cropping, can positively influence model performance, improving accuracy, recall, and F1-score. However, combining these techniques did not always lead to enhanced results, suggesting that over-processing may introduce distortions that hinder model generalization.

Data augmentation proved to be a valuable tool, with translation and rotation significantly improving model robustness and performance, particularly in recall and accuracy. Conversely, adding Gaussian noise resulted in a substantial performance decrease, indicating that it may obscure critical features in the data.

The findings emphasize the importance of careful selection and application of preprocessing and augmentation methods in medical imaging tasks. While some techniques show clear benefits, the combination of multiple preprocessing steps must be handled cautiously to avoid diminishing returns. These results underscore the need for a standardized preprocessing pipeline to enhance the reliability, reproducibility, and clinical applicability of machine learning models for schizophrenia detection.

Further work is needed to refine preprocessing pipelines and explore more advanced data augmentation methods to improve the performance of ML-based schizophrenia diagnosis. Ultimately, this study contributes to the ongoing efforts to make ML-driven medical diagnostics more robust and applicable in real-world clinical settings.

% Formalities
%\section*{Acknowledgements}

%...

\subsection*{Author contributions}

\textbf{Conception}: Prosperity Oguama, Ilia Golub; \textbf{Data Acquisition}: Ilia Golub; \textbf{Preprocessing \& Data Augmentation}: Ilia Golub; \textbf{Model Building \& Testing}: Prosperity Oguama; \textbf{Drafting \& Editing}: Ilia Golub, Prosperity Oguama. All authors contributed to the study, read, revised and approved the final manuscript.

\subsection*{Data availability statement}
The data underlying this article is publicly available and accessible through \hyperlink{Schizconnect}{http://schizconnect.org/}.

\subsection*{Financial disclosure}

The authors report no financial disclosure.

\subsection*{Conflict of interest}

The authors declare no potential conflict of interests.

\renewcommand{\baselinestretch}{1.5}
%\singlespacing
%\renewcommand\bibname{Bibliography}
\bibliographystyle{abbrv} %abbrv
\phantomsection     % 
\cleardoublepage    % 
\bibliography{paper}

\end{document}

### 2025-02-01 src/utils/augmentation.py
update aigmentation clase

@@ -1,24 +1,30 @@
 import numpy as np
 from scipy.ndimage import rotate, shift
 
-def apply_translation(image, translation=(0, 0)):
-    """Apply translation to the image along x and y axes."""
-    return shift(image, shift=(translation[0], translation[1], 0), mode='nearest')
 
-def apply_rotation(image, angle):
-    """Apply random rotation to the image."""
-    return rotate(image, angle,  axes=(0, 1),  reshape=False, mode='nearest')
-
-def apply_gaussian_noise(image, mean=0, std=0.1):
-    """
-    Adds Gaussian noise to a 3D image.
-    Args:
-        image (np.ndarray): The 3D image data.
-        mean (float): Mean of the Gaussian noise.
-        std_dev (float): Standard deviation of the Gaussian noise.
-    Returns:
-        np.ndarray: Image with added Gaussian noise.
-    """
-    std = image.std() * std
-    noise = np.random.normal(mean, std, image.shape).astype(image.dtype)
-    return image + noise
+class ImageAugmentor:
+    def __init__(self, apply_translation=True, apply_rotation=True, apply_gaussian_noise=True):
+        self.apply_translation = apply_translation
+        self.apply_rotation = apply_rotation
+        self.apply_gaussian_noise = apply_gaussian_noise
+    
+    def augment(self, image, translation=(0, 0), angle=0, mean=0, std=0.1):
+        """Applies selected augmentations to the image."""
+        if self.apply_translation:
+            image = self._apply_translation(image, translation)
+        if self.apply_rotation:
+            image = self._apply_rotation(image, angle)
+        if self.apply_gaussian_noise:
+            image = self._apply_gaussian_noise(image, mean, std)
+        return image
+    
+    def _apply_translation(self, image, translation):
+        return shift(image, shift=(translation[0], translation[1], 0), mode='nearest')
+    
+    def _apply_rotation(self, image, angle):
+        return rotate(image, angle, axes=(0, 1), reshape=False, mode='nearest')
+    
+    def _apply_gaussian_noise(self, image, mean, std):
+        std = image.std() * std
+        noise = np.random.normal(mean, std, image.shape).astype(image.dtype)
+        return image + noise

### 2025-02-01 src/augmentation_pipeline.py
Update classes

@@ -1,7 +1,6 @@
 import os
 import torch
 import random
-
 from utils.preprocess import load_nii, resample_image
 from utils.augmentation import apply_translation, apply_rotation, apply_gaussian_noise
 

### 2025-02-01 src/models/models.py

@@ -3,7 +3,6 @@ import torch.nn as nn
 from torchvision import models
 from sklearn.svm import SVC
 from sklearn.pipeline import make_pipeline
-from sklearn.model_selection import GridSearchCV
 from sklearn.preprocessing import StandardScaler
 
 # ResNet-18 for Feature Extraction
@@ -32,8 +31,22 @@ class FeatureExtractor(nn.Module):
         return x.view(x.size(0), -1)  # Flatten to (batch_size, feature_dim)
     
 # Support Vector Classifier for image classification
-def vector_classifier():
-    # Values are hardcoded after a grid search
-    svm_classifier = make_pipeline(StandardScaler(), 
-                               SVC(kernel='rbf', probability=True, C= 100, gamma= 0.0001, random_state=42))
-    return svm_classifier
+class SVMClassifier:
+    def __init__(self, kernel='rbf', C=100, gamma=0.0001, random_state=42):
+        """Initializes the SVM classifier with specified parameters."""
+        self.pipeline = make_pipeline(
+            StandardScaler(),
+            SVC(kernel=kernel, probability=True, C=C, gamma=gamma, random_state=random_state)
+        )
+    
+    def train(self, train_features, train_labels):
+        """Trains the SVM classifier on the given data."""
+        self.pipeline.fit(train_features, train_labels)
+    
+    def predict(self, test_features):
+        """Predicts labels for the given test features."""
+        return self.pipeline.predict(test_features)
+    
+    def predict_proba(self, test_features):
+        """Returns probability estimates for the given test features."""
+        return self.pipeline.predict_proba(test_features)

### 2025-02-01 src/utils/data_loader.py

@@ -3,7 +3,6 @@ import glob
 import torch
 import pandas as pd
 from torch.utils.data import Dataset
-from src.utils.preprocess import load_nii, preprocess_image
 
 class MRIDataset(Dataset):
     def __init__(self, image_paths, labels_file, target_shape=(192, 256, 256), transform=None):

### 2025-02-01 src/utils/model_plotter.py

@@ -1,0 +1,32 @@
+import matplotlib.pyplot as plt
+import seaborn as sns
+from sklearn.metrics import confusion_matrix, roc_curve, auc
+
+class SVMVisualizer:
+    def __init__(self):
+        """Initializes the visualization tools for SVM classification results."""
+        pass
+    
+    def plot_confusion_matrix(self, true_labels, predicted_labels, class_names=None):
+        """Plots the confusion matrix."""
+        cm = confusion_matrix(true_labels, predicted_labels)
+        plt.figure(figsize=(6, 5))
+        sns.heatmap(cm, annot=True, fmt='d', cmap='Blues', xticklabels=class_names, yticklabels=class_names)
+        plt.xlabel('Predicted Labels')
+        plt.ylabel('True Labels')
+        plt.title('Confusion Matrix')
+        plt.show()
+    
+    def plot_roc_curve(self, true_labels, predicted_probabilities):
+        """Plots the ROC curve."""
+        fpr, tpr, _ = roc_curve(true_labels, predicted_probabilities[:, 1])
+        roc_auc = auc(fpr, tpr)
+        
+        plt.figure(figsize=(6, 5))
+        plt.plot(fpr, tpr, color='blue', label=f'ROC curve (area = {roc_auc:.2f})')
+        plt.plot([0, 1], [0, 1], color='gray', linestyle='--')
+        plt.xlabel('False Positive Rate')
+        plt.ylabel('True Positive Rate')
+        plt.title('Receiver Operating Characteristic (ROC) Curve')
+        plt.legend()
+        plt.show()

### 2025-02-01 src/utils/augmentation.py
Add visualization class

@@ -3,11 +3,15 @@ from scipy.ndimage import rotate, shift
 
 
 class ImageAugmentor:
-    def __init__(self, apply_translation=True, apply_rotation=True, apply_gaussian_noise=True):
+
+    def __init__(self,
+                 apply_translation=True,
+                 apply_rotation=True,
+                 apply_gaussian_noise=True):
         self.apply_translation = apply_translation
         self.apply_rotation = apply_rotation
         self.apply_gaussian_noise = apply_gaussian_noise
-    
+
     def augment(self, image, translation=(0, 0), angle=0, mean=0, std=0.1):
         """Applies selected augmentations to the image."""
         if self.apply_translation:
@@ -17,13 +21,15 @@ class ImageAugmentor:
         if self.apply_gaussian_noise:
             image = self._apply_gaussian_noise(image, mean, std)
         return image
-    
+
     def _apply_translation(self, image, translation):
-        return shift(image, shift=(translation[0], translation[1], 0), mode='nearest')
-    
+        return shift(image,
+                     shift=(translation[0], translation[1], 0),
+                     mode='nearest')
+
     def _apply_rotation(self, image, angle):
         return rotate(image, angle, axes=(0, 1), reshape=False, mode='nearest')
-    
+
     def _apply_gaussian_noise(self, image, mean, std):
         std = image.std() * std
         noise = np.random.normal(mean, std, image.shape).astype(image.dtype)

### 2025-02-01 src/utils/data_visualization.py

@@ -1,0 +1,56 @@
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+class SchizophreniaEDA:
+    def __init__(self, file_path):
+        """Initializes the EDA class with the dataset."""
+        self.df = pd.read_csv(file_path)
+        self.schiz_df = self._add_schiz_column()
+        self._create_age_groups()
+    
+    def _add_schiz_column(self):
+        """Adds a binary schizophrenia column."""
+        df = self.df.copy()
+        df["schiz"] = df["dx"].apply(lambda x: "Schizophrenia" if "Schizophrenia" in x else "No Schizophrenia")
+        return df
+    
+    def _create_age_groups(self):
+        """Creates age group bins."""
+        bins = [0, 20, 30, 40, 50, 60, float('inf')]
+        labels = ["<20", "20-29", "30-39", "40-49", "50-59", "60-69"]
+        self.schiz_df["age_group"] = pd.cut(self.schiz_df["age"], bins=bins, labels=labels, right=False)
+    
+    def plot_age_distribution(self, save_path='schiz_age.png'):
+        """Plots age group distribution by schizophrenia diagnosis."""
+        age_group_dx_bin = self.schiz_df.groupby(["age_group", "schiz"], observed=False).size().unstack(fill_value=0)
+        fig = age_group_dx_bin.plot(kind='bar', stacked=False, figsize=(10, 6))
+        plt.xlabel('Age Group')
+        plt.ylabel('Count')
+        plt.xticks(rotation=45)
+        plt.legend(title='Diagnosis', bbox_to_anchor=[1,0.9])
+        plt.tight_layout()
+        plt.show()
+        fig.get_figure().savefig(save_path)
+    
+    def plot_gender_distribution(self, save_path='schiz_gen.png'):
+        """Plots schizophrenia diagnosis by gender."""
+        schiz_bin_gender = self.schiz_df.groupby(["sex", "schiz"], observed=False).size().unstack(fill_value=0)
+        fig = schiz_bin_gender.plot(kind='bar', stacked=False, colormap="Paired", figsize=(10, 6))
+        plt.xlabel('Gender')
+        plt.ylabel('Count')
+        plt.xticks(rotation=0)
+        plt.legend(title='Diagnosis', bbox_to_anchor=[1,0.9])
+        plt.tight_layout()
+        plt.show()
+        fig.get_figure().savefig(save_path)
+    
[…]

### 2025-02-01 src/main.py
add main

@@ -1,0 +1,64 @@
+import os
+import argparse
+import numpy as np
+import matplotlib.pyplot as plt
+from image_preprocessor import ImagePreprocessor
+from image_augmentor import ImageAugmentor
+from schizophrenia_eda import SchizophreniaEDA
+from svm_classifier import SVMClassifier
+from svm_visualizer import SVMVisualizer
+
+
+def main(args):
+    # Initialize components
+    preprocessor = ImagePreprocessor()
+    augmentor = ImageAugmentor()
+    eda = SchizophreniaEDA()
+    classifier = SVMClassifier()
+    visualizer = SVMVisualizer()
+
+    # Step 1: Load and preprocess images
+    print("Loading and preprocessing images...")
+    images, labels = preprocessor.load_data(args.data_path)
+    processed_images = [preprocessor.process(image) for image in images]
+
+    # Step 2: Data Augmentation (optional)
+    if args.augment:
+        print("Applying data augmentation...")
+        augmented_images, augmented_labels = augmentor.augment_data(processed_images, labels)
+        processed_images.extend(augmented_images)
+        labels.extend(augmented_labels)
+
+    # Step 3: Exploratory Data Analysis (EDA)
+    print("Performing exploratory data analysis...")
+    eda.perform_analysis(processed_images, labels)
+
+    # Step 4: Train the SVM Classifier
+    print("Training SVM classifier...")
+    X_train, X_test, y_train, y_test = classifier.split_data(processed_images, labels)
+    classifier.train(X_train, y_train)
+    accuracy, report = classifier.evaluate(X_test, y_test)
+    print(f"Model Accuracy: {accuracy:.2f}")
+    print("Classification Report:\n", report)
+
+    # Step 5: Visualize SVM results
+    print("Visualizing classifier results...")
+    visualizer.plot_decision_boundary(X_train, y_train, classifier.model)
+    visualizer.plot_confusion_matrix(y_test, classifier.predict(X_test))
+    plt.show()
+
+    # Step 6: Save trained model
+    if args.save_model:
+        classifier.save_model(args.model_path)
+        print(f"Model saved to {args.model_path}")
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(description="Schizophrenia MRI Classification Pipeline")
+    parser.add_argument("--data_path", type=str, required=True, help="Path to MRI dataset")
+    parser.add_argument("--augment", action="store_true", help="Apply data augmentation")
[…]

### 2025-02-01 src/main.py
Update main

@@ -1,12 +1,11 @@
-import os
 import argparse
-import numpy as np
 import matplotlib.pyplot as plt
-from image_preprocessor import ImagePreprocessor
-from image_augmentor import ImageAugmentor
-from schizophrenia_eda import SchizophreniaEDA
-from svm_classifier import SVMClassifier
-from svm_visualizer import SVMVisualizer
+from utils.preprocess import ImagePreprocessor
+from utils.augmentation import ImageAugmentor
+from utils.data_visualization import SchizophreniaEDA
+from models.models import SVMClassifier
+from models.models import FeatureExtractor
+from utils.model_plotter import SVMVisualizer
 
 
 def main(args):
@@ -14,7 +13,8 @@ def main(args):
     preprocessor = ImagePreprocessor()
     augmentor = ImageAugmentor()
     eda = SchizophreniaEDA()
-    classifier = SVMClassifier()
+    feature_extractor= FeatureExtractor()
+    svm_classifier = SVMClassifier()
     visualizer = SVMVisualizer()
 
     # Step 1: Load and preprocess images
@@ -32,22 +32,28 @@ def main(args):
     # Step 3: Exploratory Data Analysis (EDA)
     print("Performing exploratory data analysis...")
     eda.perform_analysis(processed_images, labels)
-
-    # Step 4: Train the SVM Classifier
+    
+    # Step 4: Extract features with ResNet-18
+    print("Extracting features...")
+    train_features, train_labels = extract_features(train_loader, feature_extractor, device)
+    
+    # Step 5: Train the SVM Classifier
     print("Training SVM classifier...")
-    X_train, X_test, y_train, y_test = classifier.split_data(processed_images, labels)
-    classifier.train(X_train, y_train)
-    accuracy, report = classifier.evaluate(X_test, y_test)
+    svm_classifier.pipeline()
+    svm_classifier.train(train_features, train_labels)
+    predicted_labels = svm_classifier.predict(test_features)
+    test_scores = svm_classifier.predict_proba(test_features)
+    
     print(f"Model Accuracy: {accuracy:.2f}")
     print("Classification Report:\n", report)
 
-    # Step 5: Visualize SVM results
+    # Step 6: Visualize SVM results
     print("Visualizing classifier results...")
-    visualizer.plot_decision_boundary(X_train, y_train, classifier.model)
-    visualizer.plot_confusion_matrix(y_test, classifier.predict(X_test))
[…]

### 2025-02-01 src/utils/model_plotter.py

@@ -17,9 +17,9 @@ class SVMVisualizer:
         plt.title('Confusion Matrix')
         plt.show()
     
-    def plot_roc_curve(self, true_labels, predicted_probabilities):
+    def plot_roc_curve(self, true_labels, test_scores):
         """Plots the ROC curve."""
-        fpr, tpr, _ = roc_curve(true_labels, predicted_probabilities[:, 1])
+        fpr, tpr, _ = roc_curve(true_labels, test_scores[:, 1])
         roc_auc = auc(fpr, tpr)
         
         plt.figure(figsize=(6, 5))

### 2025-02-01 src/utils/preprocess.py

@@ -6,372 +6,298 @@ import nibabel as nib
 from typing import Union
 from scipy.ndimage import gaussian_filter
 from skimage.filters import threshold_otsu
-from nibabel.processing import resample_to_output 
+from nibabel.processing import resample_to_output
 
-def load_nii(file_path: str) -> nib.Nifti1Image:
-    """
-    Load a nifti file from a given file path
 
-    Args:
-    file_path: str: path to the nifti file
-
-    Returns:
-    nib.Nifti1Image: nifti image object
-
-    Raises:
-    TypeError: If file_path is not a string
-    FileNotFoundError: If the file does not exist
-    ValueError: If the file is not a valid NIfTI file
-    """
-    if not isinstance(file_path, str): # Ensure the file_path is a string
-        raise TypeError(f"Expected a string for file_path, got {type(file_path)}")
+class ImagePreprocessor:
     
-    if not os.path.exists(file_path): # Ensure the file_path exists
-        raise FileNotFoundError(f"File not found: {file_path}")
+    @staticmethod
+    def load_nii(file_path: str) -> nib.Nifti1Image:
+        """
+        Load a nifti file from a given file path
 
-    try:
-        nii = nib.load(file_path)
-    except nib.filebasedimages.ImageFileError:
-        raise ValueError(f"File is not a valid NIfTI file: {file_path}")
-    except Exception as e:
-        raise RuntimeError(f"An unexpected error occurred while loading the file: {str(e)}")
+        Args:
+        file_path: str: path to the nifti file
 
-    if not isinstance(nii, nib.Nifti1Image):
-        raise ValueError(f"Loaded object is not a NIfTI image: {type(nii)}")
-
-    return nii
-
-def get_data(nii: nib.Nifti1Image) -> np.ndarray:
-    """
-    Get the data froma a nifti image object
+        Returns:
+        nib.Nifti1Image: nifti image object
+        """
+        if not isinstance(file_path, str):
+            raise TypeError(f"Expected a string for file_path, got {type(file_path)}")
     
-    Args:
-    nii: nib.Nifti1Image: nifti image object
-    
-    Returns:
-    np.ndarray: nifti image data
+        if not os.path.exists(file_path):
+            raise FileNotFoundError(f"File not found: {file_path}")
 
-    Raises:
-    TypeError: If the input is not a nib.Nifti1Image object
-    ValueError: If the image data is empty or corrupted
-    """
-    if not isinstance(nii, nib.Nifti1Image): # Ensure the image is a nibabel Nifti1Image
[…]

### 2025-02-01 src/create_train_set.py
new version

@@ -31,40 +31,3 @@ for file_path in df["Filename"]:  # Adjust column name if needed
         print(f"File not found: {src}")
 
 print("All selected files processed!")
-
-
-'''import os
-import shutil
-import pandas as pd
-
-# Paths
-csv_path = os.path.abspath("selected_files.csv")
-source_dir = os.path.abspath("data/processed/extracted_brain")
-destination_dir = os.path.abspath("data/processed/extracted_brain/train_set")
-
-# Ensure destination directory exists
-os.makedirs(destination_dir, exist_ok=True)
-
-# Load the CSV file
-df = pd.read_csv(csv_path)
-
-# Move files
-counter = 0
-for file_name in df["Filename"].str.strip():  # Assuming CSV has a column 'filename'
-    file_name = file_name.strip()
-    src = os.path.join(source_dir, file_name)
-    dst = os.path.join(destination_dir, file_name)
-
-    print(f"Checking file: {repr(file_name)}")  # Debugging step
-    print(f"Source path: {src}")
-    print(f"Destination path: {dst}")
-
-    if os.path.exists(src):
-        shutil.move(src, dst)
-        print(f"Moved: {file_name}")
-        counter += 1
-        print(f"Moved {counter} file(s).")
-    else:
-        print(f"File not found: {file_name}")
-
-print("All selected files moved successfully!")'''

### 2025-02-02 src/paper/paper.tex
temp

@@
 This was done to ensure uniformity of the data obtained.
+Figure~\ref{fig:raw_mri_volume_slices} presents the example of raw data.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{./figs/sample_volume.png} % test_pipeline.png
+    \caption{Three image slices representing the average of a 3D column in three views: dorsal, lateral, and superior respectively}\label{fig:raw_mri_volume_slices}
+\end{figure}
 

### 2025-02-02 src/augmentation_pipeline.py

@@ -1,86 +1,0 @@
-import os
-import torch
-import random
-from utils.preprocess import load_nii, resample_image
-from utils.augmentation import apply_translation, apply_rotation, apply_gaussian_noise
-
-def preprocess_and_augment(file_path, augmentation_type, num_augmentations=1, 
-                           translation_range=(-10, 10), rotation_degrees=(-20, 20), noise_mean=0, noise_std=0.1):
-    """Apply augmentation (translation, rotation, or gaussian noise) to the image."""
-    nii = load_nii(file_path)
-    image_data = resample_image(nii, voxel_size=(2, 2, 2), order=4, mode='reflect', cval=0)
-    #print(f"Original shape: {image_data.shape}")  # Debugging step
-    
-    augmented_images = []
-
-    for _ in range(num_augmentations):
-        if augmentation_type == 'translation':
-            translation = (random.randint(translation_range[0], translation_range[1]), 
-                           random.randint(translation_range[0], translation_range[1]))  # Random translation in x and y axes
-            augmented_image = apply_translation(image_data, translation)
-        elif augmentation_type == 'rotation':
-            angle = random.randint(rotation_degrees[0], rotation_degrees[1])  # Random rotation angle between specified degrees
-            augmented_image = apply_rotation(image_data, angle)
-        elif augmentation_type == 'gaussian_noise':
-            augmented_image = apply_gaussian_noise(image_data, mean=noise_mean, std=noise_std)
-        else:
-            raise ValueError(f"Unknown augmentation type: {augmentation_type}")
-        
-        #print(f"Augmented shape: {augmented_image.shape}")  # Debugging step
-
-        augmented_images.append(augmented_image)
-    
-    return augmented_images
-
-def save_as_tensor(image_data, save_dir, filename):
-    """Save the augmented image as a PyTorch tensor."""
-    tensor_image = torch.tensor(image_data, dtype=torch.float32) # , dtype=torch.float32
-    save_path = os.path.join(save_dir, f"{filename}.pt")
-    torch.save(tensor_image, save_path)
-    #print(f"Saved tensor shape: {tensor_image.shape}")  # Debugging
-    print(f"Saved augmented image: {save_path}")
-
-def augment_images_in_directory(directory, augmentation_type, num_augmentations=1, 
-                                 translation_range=(-10, 10), rotation_degrees=(0, 360), noise_mean=0, noise_std=0.1):
-    """Iterate over all .nii.gz files in a directory and apply augmentations."""
-    output_dir = 'data/processed/augmented'
[…]

### 2025-02-02 src/best_pipeline.py

@@ -1,58 +1,0 @@
-import os
-import torch
-import random
-import numpy as np
-import pandas as pd
-from utils.augmentation import apply_translation, apply_rotation
-
-def save_as_tensor(image_data, save_dir, filename):
-    """Save the augmented image as a PyTorch tensor."""
-    os.makedirs(save_dir, exist_ok=True)
-    tensor_image = torch.tensor(image_data, dtype=torch.float32)
-    save_path = os.path.join(save_dir, f"{filename}.pt")
-    torch.save(tensor_image, save_path)
-    print(f"Saved: {save_path}")
-
-def augment_and_save(file_path, save_dir, augmentation_type, num_augmentations, value_range):
-    """Applies augmentations and saves the output."""
-    image_data = torch.load(file_path).numpy()
-    filename = os.path.splitext(os.path.basename(file_path))[0]
-    
-    for i in range(1, num_augmentations + 1):
-        if augmentation_type == "rotation":
-            angle = random.uniform(*value_range)
-            augmented_image = apply_rotation(image_data, angle)
-        elif augmentation_type == "translation":
-            translation = (random.randint(*value_range), random.randint(*value_range))
-            augmented_image = apply_translation(image_data, translation)
-        else:
-            raise ValueError("Invalid augmentation type")
-        
-        save_as_tensor(augmented_image, save_dir, f"{filename}_{augmentation_type}_{i}")
-
-def main():
-    input_dir = "data/processed/extracted_brain"
-    output_dir = "data/fully_processed/augmented"
-    num_files = 50  # Number of randomly selected files
-    num_augmentations = 3
-    rotation_range = (-10, 10)
-    translation_range = (-10, 10)
-    
-    # Get all .pt files and randomly select 50
-    all_files = [os.path.join(input_dir, f) for f in os.listdir(input_dir) if f.endswith(".pt")]
-    selected_files = random.sample(all_files, num_files)
-    
-    # Save selected filenames to a CSV file
-    pd.DataFrame(selected_files, columns=["Filename"]).to_csv("selected_files.csv", index=False)
-    print("Saved selected filenames to selected_files.csv")
-    
-    # Apply augmentations
-    counter = 0
-    for file_path in selected_files:
-        augment_and_save(file_path, os.path.join(output_dir, "rotated"), "rotation", num_augmentations, rotation_range)
-        augment_and_save(file_path, os.path.join(output_dir, "translated"), "translation", num_augmentations, translation_range)
[…]

### 2025-02-02 src/create_raw_train_set.py

@@ -1,26 +1,0 @@
-import os
-import shutil
-import pandas as pd
-
-# Paths
-csv_file = "selected_files.csv"
-src_dir = "data/raw_pt"
-dst_dir = "data/raw_pt/basis_for_train_set"
-
-# Ensure destination directory exists
-os.makedirs(dst_dir, exist_ok=True)
-
-# Load CSV and extract filenames
-df = pd.read_csv(csv_file)
-df["Filename"] = df["Filename"].apply(lambda x: os.path.basename(x))  # Extract only filename
-
-# Move files
-for filename in df["Filename"]:
-    src_path = os.path.join(src_dir, filename)
-    dst_path = os.path.join(dst_dir, filename)
-
-    if os.path.exists(src_path):
-        shutil.move(src_path, dst_path)
-        print(f"Moved: {filename}")
-    else:
-        print(f"File not found: {filename}")

### 2025-02-02 src/create_train_set.py

@@ -1,33 +1,0 @@
-import os
-import shutil
-import pandas as pd
-
-# Get absolute paths
-csv_path = os.path.abspath("selected_files.csv")
-destination_dir = os.path.abspath("data/processed/extracted_brain/train_set")
-
-# Ensure destination directory exists
-os.makedirs(destination_dir, exist_ok=True)
-
-# Load the CSV file
-df = pd.read_csv(csv_path)
-
-# Move files
-for file_path in df["Filename"]:  # Adjust column name if needed
-    file_path = file_path.strip()  # Remove any unwanted spaces
-
-    # Convert relative path from CSV to absolute path
-    src = os.path.abspath(file_path)
-    dst = os.path.join(destination_dir, os.path.basename(file_path))
-
-    print(f"Checking file: {repr(file_path)}")  # Debugging step
-    print(f"Source path: {src}")
-    print(f"Destination path: {dst}")
-
-    if os.path.exists(src):
-        shutil.move(src, dst)
-        print(f"Moved: {file_path}")
-    else:
-        print(f"File not found: {src}")
-
-print("All selected files processed!")

### 2025-02-02 src/preprocessing_pipeline.py

@@ -1,169 +1,0 @@
-import os
-import torch
-import nibabel as nib
-import matplotlib.pyplot as plt
-
-from utils.preprocess import (
-    load_nii,
-    get_data,
-    resample_image,
-    normalize_data,
-    extract_brain,
-    crop_to_largest_bounding_box,
-    apply_gaussian_smoothing
-)
-
-from utils.preprocess_validation import calculate_metrics
-
-def preprocess_pipeline(file_path, 
-                        steps_to_take: list, 
-                        what_to_return: list, 
-                        separation: bool = False):
-    """
-    Preprocessing pipeline for NIfTI files.
-    
-    Args:
-        file_path (str): Path to the .nii.gz file.
-        steps_to_take (list): Steps to apply during preprocessing.
-        what_to_return (list): Outputs to save and return.
-    
-    Returns:
-        dict: Dictionary with processed outputs requested in `what_to_return`.
-    """
-    # Dictionary to store the output
-    to_be_returned = {}
-
-    # Load NIfTI file
-    nii = load_nii(file_path)
-
-    # Resample image
-    if 'resample' in steps_to_take :
-        resampled_image = resample_image(nii, voxel_size=(2, 2, 2), order=4, mode='reflect', cval=0)
-    else:
-        pass
-    
-    if 'resampled_image' in what_to_return:
-        to_be_returned['resampled_image'] = resampled_image
-
-    # Normalize data
-    if 'normalize' in steps_to_take:
-        normalized_image = normalize_data(resampled_image)
-    else:
-        pass
-    
-    if 'normalized_image' in what_to_return:
-        to_be_returned['normalized_image'] = normalized_image
-
-    # Brain extraction
-    if 'extract_brain' in steps_to_take:
-        if separation:
-            brain_data = extract_brain(resampled_image, what_to_return={'extracted_brain': 'numpy', 'mask': 'numpy'})
-        else:
-            brain_data = extract_brain(normalized_image, what_to_return={'extracted_brain': 'numpy', 'mask': 'numpy'})
-        extracted_brain = brain_data['extracted_brain']
-        mask = brain_data['mask']
-    else:
-        mask = None
-    
-    if 'extracted_brain' in what_to_return:
-        to_be_returned['extracted_brain'] = extracted_brain
-
-    # Cropping
-    if 'crop' in steps_to_take:
-        if separation:
-            brain_data = extract_brain(resampled_image, what_to_return={'mask': 'numpy'})
-            mask = brain_data['mask']
[…]

### 2025-02-02 src/main.py
clean workflow

@@ -1,70 +1,29 @@
-import argparse
-import matplotlib.pyplot as plt
-from utils.preprocess import ImagePreprocessor
-from utils.augmentation import ImageAugmentor
-from utils.data_visualization import SchizophreniaEDA
-from models.models import SVMClassifier
-from models.models import FeatureExtractor
-from utils.model_plotter import SVMVisualizer
+import os
+from preprocessing import preprocess_images
+from augmentation import augment_images
+from dataset_preparation import prepare_dataset
 
-
-def main(args):
-    # Initialize components
-    preprocessor = ImagePreprocessor()
-    augmentor = ImageAugmentor()
-    eda = SchizophreniaEDA()
-    feature_extractor= FeatureExtractor()
-    svm_classifier = SVMClassifier()
-    visualizer = SVMVisualizer()
-
-    # Step 1: Load and preprocess images
-    print("Loading and preprocessing images...")
-    images, labels = preprocessor.load_data(args.data_path)
-    processed_images = [preprocessor.process(image) for image in images]
-
-    # Step 2: Data Augmentation (optional)
-    if args.augment:
-        print("Applying data augmentation...")
-        augmented_images, augmented_labels = augmentor.augment_data(processed_images, labels)
-        processed_images.extend(augmented_images)
-        labels.extend(augmented_labels)
-
-    # Step 3: Exploratory Data Analysis (EDA)
-    print("Performing exploratory data analysis...")
-    eda.perform_analysis(processed_images, labels)
+def main():
+    # Define paths
+    raw_data_dir = "data/raw"
+    preprocessed_dir = "data/processed"
+    augmented_dir = "data/fully_processed"
+    train_set_dir = "data/train_set"
     
-    # Step 4: Extract features with ResNet-18
-    print("Extracting features...")
-    train_features, train_labels = extract_features(train_loader, feature_extractor, device)
-    
-    # Step 5: Train the SVM Classifier
-    print("Training SVM classifier...")
-    svm_classifier.pipeline()
-    svm_classifier.train(train_features, train_labels)
-    predicted_labels = svm_classifier.predict(test_features)
-    test_scores = svm_classifier.predict_proba(test_features)
-    
-    print(f"Model Accuracy: {accuracy:.2f}")
-    print("Classification Report:\n", report)
+    # User Input
+    num_train_images = 50
+    preprocessing_steps = ["resample", "normalize", "extract_brain"]
[…]

### 2025-02-02 src/utils/dataset_preparation.py

@@ -1,0 +1,20 @@
+import os
+import random
+import shutil
+import pandas as pd
+
+def prepare_dataset(input_dir, train_dir, num_train_images):
+    os.makedirs(train_dir, exist_ok=True)
+
+    all_files = [f for f in os.listdir(input_dir) if f.endswith(".pt")]
+    selected_files = random.sample(all_files, min(num_train_images, len(all_files)))
+
+    # Save selected filenames
+    df = pd.DataFrame(selected_files, columns=["Filename"])
+    df.to_csv("selected_files.csv", index=False)
+
+    for file in selected_files:
+        src = os.path.join(input_dir, file)
+        dst = os.path.join(train_dir, file)
+        shutil.copy(src, dst)
+        print(f"Copied {file} to train set")

### 2025-02-02 src/main.py
minor

@@ -1,7 +1,6 @@
-import os
-from preprocessing import preprocess_images
-from augmentation import augment_images
-from dataset_preparation import prepare_dataset
+from utils.preprocessing import preprocess_images
+from utils.augmentation import augment_images
+from utils.dataset_preparation import prepare_dataset
 
 def main():
     # Define paths

### 2025-02-02 src/utils/preprocess.py
restore preprocess.py

@@ -6,298 +6,372 @@ import nibabel as nib
 from typing import Union
 from scipy.ndimage import gaussian_filter
 from skimage.filters import threshold_otsu
-from nibabel.processing import resample_to_output
+from nibabel.processing import resample_to_output 
 
+def load_nii(file_path: str) -> nib.Nifti1Image:
+    """
+    Load a nifti file from a given file path
 
-class ImagePreprocessor:
+    Args:
+    file_path: str: path to the nifti file
+
+    Returns:
+    nib.Nifti1Image: nifti image object
+
+    Raises:
+    TypeError: If file_path is not a string
+    FileNotFoundError: If the file does not exist
+    ValueError: If the file is not a valid NIfTI file
+    """
+    if not isinstance(file_path, str): # Ensure the file_path is a string
+        raise TypeError(f"Expected a string for file_path, got {type(file_path)}")
     
-    @staticmethod
-    def load_nii(file_path: str) -> nib.Nifti1Image:
-        """
-        Load a nifti file from a given file path
+    if not os.path.exists(file_path): # Ensure the file_path exists
+        raise FileNotFoundError(f"File not found: {file_path}")
 
-        Args:
-        file_path: str: path to the nifti file
+    try:
+        nii = nib.load(file_path)
+    except nib.filebasedimages.ImageFileError:
+        raise ValueError(f"File is not a valid NIfTI file: {file_path}")
+    except Exception as e:
+        raise RuntimeError(f"An unexpected error occurred while loading the file: {str(e)}")
 
-        Returns:
-        nib.Nifti1Image: nifti image object
-        """
-        if not isinstance(file_path, str):
-            raise TypeError(f"Expected a string for file_path, got {type(file_path)}")
+    if not isinstance(nii, nib.Nifti1Image):
+        raise ValueError(f"Loaded object is not a NIfTI image: {type(nii)}")
+
+    return nii
+
+def get_data(nii: nib.Nifti1Image) -> np.ndarray:
+    """
+    Get the data froma a nifti image object
     
-        if not os.path.exists(file_path):
-            raise FileNotFoundError(f"File not found: {file_path}")
+    Args:
+    nii: nib.Nifti1Image: nifti image object
+    
+    Returns:
+    np.ndarray: nifti image data
 
-        try:
-            nii = nib.load(file_path)
-        except nib.filebasedimages.ImageFileError:
-            raise ValueError(f"File is not a valid NIfTI file: {file_path}")
+    Raises:
[…]

### 2025-02-02 src/paper/paper.tex
updated paper

@@
 Table~\ref{tab:preprocessing_pipeline} presents the preprocessing techniques used and their respective hyperparameters and tools used.
-Figure~\ref{fig:test_pipeline} gives a visual representation of the conducted tests.
+Figure~\ref{fig:test_pipeline} gives a comparison of a raw image and a preprocessed image, also indicating the image metrics.
+%Figure~\ref{fig:test_pipeline} gives a visual representation of the conducted tests.
 %
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{./figs/empty.pdf} % test_pipeline.png
-    \caption{Tests pipeline for MRI data.}\label{fig:test_pipeline}
+    \includegraphics[width=0.4\textwidth]{./figs/slice_comparison.png} % test_pipeline.png
+    \caption{Panel (A) shows a slice of a raw scan.
+Panel (B) shows a slice of a preprocessed scan.
+For preprocessed scans, the standard deviation of SNR values is relatively low (\mbox{$SNR= 25 \pm 1$}), which ensures consistency in the image quality and prevents the model from making predictions based on possible artifacts.}\label{fig:test_pipeline}
 \end{figure}
@@
 Cross validation was performed with the support vector classifier (SVC) to choose the best hyperparameters, and the hyperparameters that yielded the best results were the radial basis function (RBF) kernel, a gamma value of 0.0001 and a C parameter value of 100.
-Each 2D slice of the images were fed into the ResNet-18 model as NumPy arrays of dimension $224x224$.
+Each 2D slice of the images were fed into the ResNet-18 model as NumPy arrays of dimension \texttt{224x224}.
 This input size corresponds to the size of the images in the ImageNet dataset that was used to train ResNet-18 and was selected for compatibility with the pretrained weights of the ResNet-18 model.

### 2025-02-02 tests/test_data_loader.py
minor changes

@@ -5,11 +5,11 @@ import pandas as pd
 import glob
 from unittest.mock import patch, MagicMock
 from utils.preprocess import load_nii
-from utils import data_loader 
+from utils.data_loader import MRIDataset 
 
 def mock_load_nii(file_path):
     """Mock function to simulate loading a NIfTI image."""
-    return torch.rand((1, 128, 128, 128))  # Simulating a 3D tensor
+    return load_nii(torch.rand((1, 128, 128, 128)))  # Simulating a 3D tensor
 
 def mock_preprocess_image(image):
     """Mock function to simulate preprocessing an image."""

### 2025-02-02 tests/test_preprocess.py

@@ -6,21 +6,16 @@ import numpy as np
 from unittest.mock import patch
 from tempfile import NamedTemporaryFile
 from unittest.mock import patch, MagicMock
-from scipy.ndimage import binary_closing, binary_opening
-from skimage.filters import threshold_otsu
-from scipy.ndimage import gaussian_filter
-
-
 
 # Import functions
-from utils import load_nii 
-from utils import get_data
-from utils import get_affine
-from utils import match_dimensions
-from utils import resample_image
-from utils import normalize_data, crop_to_largest_bounding_box, get_largest_brain_mask_slice
-from utils import refine_brain_mask, get_largest_brain_mask_slice
-from utils import apply_gaussian_smoothing
+from utils.preprocess import load_nii 
+from utils.preprocess import get_data
+from utils.preprocess import get_affine
+from utils.preprocess import match_dimensions
+from utils.preprocess import resample_image
+from utils.preprocess import normalize_data, crop_to_largest_bounding_box, get_largest_brain_mask_slice
+from utils.preprocess import get_largest_brain_mask_slice
+from utils.preprocess import apply_gaussian_smoothing
 
 
 class TestLoadNii(unittest.TestCase):
@@ -379,78 +374,6 @@ class TestNormalizeData(unittest.TestCase):
         # Ensure that the function still works with a small epsilon even if there's zero variation
         self.assertEqual(np.min(result), 0)
         self.assertEqual(np.max(result), 0)
-
-
-class TestRefineBrainMask(unittest.TestCase):
-
-    def test_refine_brain_mask(self):
-        # Create a simple binary mask
-        mask = np.array([[0, 0, 1, 1, 0],
-                         [0, 1, 1, 1, 0],
-                         [0, 1, 1, 1, 0],
-                         [0, 0, 1, 1, 0],
-                         [0, 0, 0, 0, 0]], dtype=np.uint8)
-        
-        # Call refine_brain_mask with no structure provided
-        refined_mask = refine_brain_mask(mask)
-
-        # Check that the result is a numpy array
-        self.assertIsInstance(refined_mask, np.ndarray)
-
-        # Check if the result contains only 0s and 1s (binary mask)
-        self.assertTrue(np.all(np.isin(refined_mask, [0, 1])))
-
-        # Check that the refined mask looks as expected (after closing and opening)
-        expected_refined_mask = binary_opening(binary_closing(mask), structure=None)
[…]

### 2025-02-02 tests/test_data_loader.py
fix test

@@ -15,8 +15,8 @@ def mock_preprocess_image(image):
     """Mock function to simulate preprocessing an image."""
     return image  # Assume it returns the same tensor
 
-@patch("utils.preprocess.load_nii", side_effect=mock_load_nii)
-@patch("dataset.preprocess_image", side_effect=mock_preprocess_image)
+@patch("src.utils.data_loader", side_effect=mock_load_nii)
+@patch("src.utils.preprocess", side_effect=mock_preprocess_image)
 @patch("pandas.read_csv")
 @patch("glob.glob")
 def test_mri_dataset(mock_glob, mock_read_csv, mock_preprocess, mock_load_nii):

### 2025-02-06 src/main.py
minor updates

@@ -1,8 +1,11 @@
+from logger import logger
+
 from utils.preprocessing import preprocess_images
 from utils.augmentation import augment_images
 from utils.dataset_preparation import prepare_dataset
 
 def main():
+    logger.info("The program started...")
     # Define paths
     raw_data_dir = "data/raw"
     preprocessed_dir = "data/processed"
@@ -24,5 +27,7 @@ def main():
     # Step 3: Dataset Preparation
     prepare_dataset(augmented_dir, train_set_dir, num_train_images)
 
+    logger.info("The program completed successfully.")
+    
 if __name__ == "__main__":
     main()

### 2025-02-06 src/main.py
plan

@@ -1,4 +1,5 @@
 from logger import logger
+import yaml
 
 from utils.preprocessing import preprocess_images
 from utils.augmentation import augment_images
@@ -6,6 +7,15 @@ from utils.dataset_preparation import prepare_dataset
 
 def main():
     logger.info("The program started...")
+
+    # Load configuration from a config file
+    with open('config.yaml') as c:
+        config_data = yaml.safe_load(c)
+    
+    # Extract the necessary configuration parameters
+    # raw_data_dir = config_data['raw_data_dir']
+    #
+
     # Define paths
     raw_data_dir = "data/raw"
     preprocessed_dir = "data/processed"
@@ -18,16 +28,58 @@ def main():
     augmentation_type = "translation"  # Options: 'translation', 'rotation', 'gaussian_noise'
     num_augmentations = 5
 
-    # Step 1: Preprocessing
-    preprocess_images(raw_data_dir, preprocessed_dir, preprocessing_steps)
-
-    # Step 2: Augmentation
-    augment_images(preprocessed_dir, augmented_dir, augmentation_type, num_augmentations)
-
-    # Step 3: Dataset Preparation
+    # Step 1: Select the data (train, test)
+    """
+    Randomly divides (ratio is user defined) images located in /data/raw,
+    Creates 2 folders: train_set, test_set,
+    Copy-pastes previously divided images into these folders (e.g. 50 images into train, 12 images into test),
+    Original images are kept intact in /data/raw.
+    """
     prepare_dataset(augmented_dir, train_set_dir, num_train_images)
+    logger.info("Dataset preparation completed.")
+
+    # Step 2: Preprocessing
+    """
+    Preprocesses images located in /data/train_set,
+    Resamples to a common voxel size,
+    Normalizes intensity values,
+    Extracts brain regions using a custom algorithm,
+    Saves preprocessed images in /data/train_set/preprocessed.
+
+    Test images located in /data/test_set are also resampled and normalized for consistency,
+    Saves preprocessed test images in /data/test_set/preprocessed.
+    """
+    preprocess_images(raw_data_dir, preprocessed_dir, preprocessing_steps)
+    logger.info("Preprocessing completed.")
+
+    # Step 3: Augmentation
+    """
+    Applies specified augmentation type (translation, rotation, gaussian_noise, or combined)
+    to the preprocessed images located in /data/train_set/preprocessed,
+    Saves augmented images in /data/train_set/preprocessed/augmented.
+    """
[…]

### 2025-02-06 tests/test_preprocess.py
to fix

@@ -8,7 +8,7 @@ from tempfile import NamedTemporaryFile
 from unittest.mock import patch, MagicMock
 
 # Import functions
-from utils.preprocess import load_nii 
+from ..src.utils.preprocess import load_nii 
 from utils.preprocess import get_data
 from utils.preprocess import get_affine
 from utils.preprocess import match_dimensions

### 2025-02-06 src/main.py
to be fixed

@@ -1,6 +1,7 @@
 from logger import logger
 import yaml
 
+from utils.eda import SchizophreniaEDA
 from utils.preprocessing import preprocess_images
 from utils.augmentation import augment_images
 from utils.dataset_preparation import prepare_dataset
@@ -27,6 +28,15 @@ def main():
     preprocessing_steps = ["resample", "normalize", "extract_brain"]
     augmentation_type = "translation"  # Options: 'translation', 'rotation', 'gaussian_noise'
     num_augmentations = 5
+
+    # Step 0: Explanatory data analysis
+    """
+    Performs an explanatory data analysis on the raw images and corresponding clinical data.
+    This step is optional and can be skipped if not necessary.
+    """
+    eda = SchizophreniaEDA(PATH_TO_CLINICAL_DATA, PATH_TO_RAW_IMAGES)
+    eda.perform_eda()
+    logger.info("EDA completed.")
 
     # Step 1: Select the data (train, test)
     """

### 2025-02-06 src/utils/eda.py
fixing

@@ -2,13 +2,18 @@ import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
 import seaborn as sns
+import torch
+
+from utils.preprocess_validation import calculate_metrics
 
 class SchizophreniaEDA:
-    def __init__(self, file_path):
+    def __init__(self, path_to_clinical_data = None, path_to_raw_images = None):
         """Initializes the EDA class with the dataset."""
-        self.df = pd.read_csv(file_path)
+        self.df = pd.read_csv(path_to_clinical_data)
         self.schiz_df = self._add_schiz_column()
         self._create_age_groups()
+
+        self.path_to_raw_images = path_to_raw_images
     
     def _add_schiz_column(self):
         """Adds a binary schizophrenia column."""
@@ -54,3 +59,21 @@ class SchizophreniaEDA:
             "Schizophrenia": df_schiz["age"].describe(),
             "No Schizophrenia": df_ns["age"].describe()
         }
+    
+    def assess_raw_images(self):
+        """Assesses the quality of raw images."""
+        raw_metrics_all_images = {} # TODO: should contain paths to images
+
+        # iterate over all .pt files in path_to_raw_images
+        for path in self.path_to_raw_images:
+            loaded_tensor = torch.load(path)
+            tensor_to_numpy = loaded_tensor.numpy()
+            raw_metrics_per_image = calculate_metrics(None, tensor_to_numpy)
+            raw_metrics_all_images.append(raw_metrics_per_image)
+        return raw_metrics_all_images
+    
+    def save_raw_images_metrics(self):
+        """Saves raw image metrics to a CSV file."""
+        raw_metrics_all_images = self.assess_raw_images()
+        df_raw_metrics = pd.DataFrame(raw_metrics_all_images)
+        df_raw_metrics.to_csv('raw_image_metrics.csv', index=False)

### 2025-02-06 src/utils/preprocess_validation.py

@@ -495,8 +495,8 @@ def calculate_ssim(data1: np.ndarray, data2: np.ndarray) -> float:
     return ssim(data1, data2, data_range=data1.max() - data1.min())
 
 
-def calculate_metrics(reference_data: np.ndarray, 
-                      modified_data: np.ndarray, 
+def calculate_metrics(reference_data: np.ndarray = None, 
+                      modified_data: np.ndarray = None, 
                       preprocessing_step: str = None
                       ) -> dict:
     """
@@ -516,14 +516,19 @@ def calculate_metrics(reference_data: np.ndarray,
     # Dictionary to store the metrics values
     results_metrics = []
     # Compute metrics
-    modified_data_matched = match_dimensions(reference_data, modified_data)
+    if reference_data is not None:
+        modified_data_matched = match_dimensions(reference_data, modified_data)
 
+        mse_value = calculate_mse(reference_data, modified_data_matched)
+        psnr_value = calculate_psnr(reference_data, modified_data_matched, mse=mse_value)
+        ssim_value = calculate_ssim(reference_data, modified_data_matched)
+    else:
+        mse_value = -1
+        psnr_value = -1
+        ssim_value = -1
+        
     snr_value = calculate_snr_with_mask(modified_data)
     cnr_value = calculate_cnr(modified_data)
-
-    mse_value = calculate_mse(reference_data, modified_data_matched)
-    psnr_value = calculate_psnr(reference_data, modified_data_matched, mse=mse_value)
-    ssim_value = calculate_ssim(reference_data, modified_data_matched)
 
     relative_psnr_value = calculate_relative_psnr(modified_data)
     relative_rmse_value = calculate_relative_rmse(modified_data)

### 2025-02-13 src/utils/data_loader.py
update data_loader.py

@@ -1,75 +1,131 @@
 import re
 import torch
+import logging
+import pandas as pd
 import torch.nn.functional as F
 from torch.utils.data import Dataset
+from pathlib import Path
+from typing import Tuple, List, Optional
 
-# Function to extract subject number from the file name
-def extract_subject_number(file_name):
-    match = re.search(r'sub-(A\d+)', file_name)
-    if match:
-        return match.group(1)
-    return None
 
 # Class to return each slice and the label
 class MRIDataset(Dataset):
-    def __init__(self, image_paths, labels_file, target_shape=(128, 128)):
+    """
+    A PyTorch Dataset for loading MRI slices and their corresponding labels.
+
+    This dataset loads MRI scans from tensor files, extracts 2D slices, and resizes them.
+    Labels are matched using a provided CSV file that contains participant IDs.
+
+    Attributes:
+        image_paths (List[Path]): List of file paths to MRI tensor files.
+        labels_df (pd.DataFrame): DataFrame containing participant IDs and corresponding labels.
+        target_shape (Tuple[int, int]): Desired shape (height, width) for each 2D slice.
+    """
+
+    def __init__(
+        self,
+        image_paths: List[str],
+        labels_file: pd.DataFrame,
+        target_shape: Tuple[int, int] = (128, 128)
+    ) -> None:
         """
-        Args:
-            image_paths (list): List of file paths to MRI tensors.
-            labels_file (pd.DataFrame): DataFrame containing participant IDs and labels.
-            target_shape (tuple): Target height and width for 2D slices.
-        """
-        self.image_paths = image_paths
+            Initializes the dataset.
+
+            Parameters:
+                image_paths (List[str]): List of file paths to MRI tensors.
+                labels_file (pd.DataFrame): DataFrame containing participant IDs and labels.
+                target_shape (Tuple[int, int]): Target height and width for 2D slices.
+                
+            Returns:
+                None
+            """
+        self.image_paths = [Path(p)
+                            for p in image_paths]  # Convert to Path objects
         self.labels_df = labels_file
         self.target_shape = target_shape
 
-    def __len__(self):
-        """Calculates the total number of slices across all MRI files."""
+    def __len__(self) -> int:
+        """
+            Returns the total number of slices across all MRI files.
+
+            Returns:
+                int: Total number of 2D slices.
+            """
         total_slices = 0
         for path in self.image_paths:
             mri_tensor = torch.load(path, weights_only=True)
[…]

### 2025-02-13 src/utils/preprocess_validation.py

@@ -1,5 +1,4 @@
 from typing import Union
-
 import matplotlib.pyplot as plt
 import nibabel as nib
 import numpy as np
@@ -10,13 +9,11 @@ from skimage.metrics import structural_similarity as ssim
 from src.utils.preprocess import get_data, match_dimensions
 
 
-def plot_slices(
-    data: Union[np.ndarray, nib.Nifti1Image],
-    how_many: int = 4,
-    title: str = "",
-    axes: list = None,
-    isSave_fig: bool = False
-    ) -> None:
+def plot_slices(data: Union[np.ndarray, nib.Nifti1Image],
+                how_many: int = 4,
+                title: str = "",
+                axes: list = None,
+                isSave_fig: bool = False) -> None:
     """
     Plot N evenly spaced slices along the z-axis (axis 2) of the MRI volume (excluding the edge cases).
     Plot slices on provided axes or create a new figure if axes are not provided.
@@ -32,21 +29,23 @@ def plot_slices(
     if not isinstance(data, (np.ndarray, nib.Nifti1Image)):
         raise TypeError(
             f"Input data must be a numpy array or a NIfTI image, got {type(data)}"
-            )
+        )
 
     if data.ndim != 3:  # Ensure data is 3 dimensional
-        raise ValueError(f"Input data must be a 3D numpy array, got {data.ndim}")
+        raise ValueError(
+            f"Input data must be a 3D numpy array, got {data.ndim}")
 
     if how_many < 1 or how_many * 2 > data.shape[2]:
         raise ValueError(
             f"Number of slices to plot must be between 1 and the total number of slices, got {how_many}"
-            )
+        )
 
     try:
         if not isinstance(data, np.ndarray):
             data = get_data(data)
         z_dim = data.shape[2]  # Size along the z-axis
-        slice_indices = np.linspace(0, z_dim - 1, how_many * 2, dtype=int)  # Select evenly spaced slices
+        slice_indices = np.linspace(0, z_dim - 1, how_many * 2,
+                                    dtype=int)  # Select evenly spaced slices
 
         # Handle both odd and even values of how_many
         start_index = (len(slice_indices) - how_many) // 2
@@ -54,7 +53,8 @@ def plot_slices(
         slice_indices = slice_indices[start_index:end_index]
 
         if len(slice_indices) != how_many:
[…]

### 2025-02-13 src/utils/dataset_preparation.py
update dataset_preparation.py

@@ -2,58 +2,79 @@ import os
 import shutil
 import random
 import torch
+from typing import Optional
 from src.utils.preprocess import load_nii, resample_image, normalize_data
 
-def prepare_dataset(
-        train_ratio=0.8, 
-        raw_pt_dir="data/raw_pt", 
-        raw_nii_dir="data/raw_nii", 
-        normalize=True, 
-        norm_method="min-max"
-        ) -> None:
+
+def prepare_dataset(train_ratio: float = 0.8,
+                    raw_pt_dir: str = "data/raw_pt",
+                    raw_nii_dir: str = "data/raw_nii",
+                    normalize: bool = True,
+                    norm_method: str = "min-max") -> None:
     """
-    Randomly splits images from raw_pt_dir into train and test sets based on train_ratio.
-    Copies images into train_set and test_set for raw_pt and ensures the same split for raw_nii,
-    while also resampling and normalizing raw_nii images before saving them in .pt format.
-    """
-    # Create train/test directories for raw_pt
-    train_pt_dir = os.path.join(raw_pt_dir, "train_set")
-    test_pt_dir = os.path.join(raw_pt_dir, "test_set")
-    os.makedirs(train_pt_dir, exist_ok=True)
-    os.makedirs(test_pt_dir, exist_ok=True)
+    Splits MRI images into training and test sets while maintaining the same split across .pt and .nii formats.
     
-    # Get list of .pt files in raw_pt_dir
+    The .pt files are copied into respective train/test folders for first training experiment, while
+    .nii files are preprocessed (resampled, optionally normalized) before being saved in .pt format.
+
+    Parameters:
+        train_ratio (float): Proportion of data to use for training.
+        raw_pt_dir (str): Directory containing raw .pt images.
+        raw_nii_dir (str): Directory containing raw .nii images.
+        normalize (bool): Whether to normalize the .nii images.
+        norm_method (str): Normalization method ("min-max" or other supported methods).
+    """
+    # Define train/test directories
+    train_pt_dir, test_pt_dir = os.path.join(raw_pt_dir,
+                                             "train_set"), os.path.join(
+                                                 raw_pt_dir, "test_set")
+    train_nii_dir, test_nii_dir = os.path.join(raw_nii_dir,
+                                               "train_set"), os.path.join(
+                                                   raw_nii_dir, "test_set")
+
[…]

### 2025-02-14 src/main.py
updated workflow

@@ -3,9 +3,9 @@ from omegaconf import OmegaConf
 from utils.eda import SchizophreniaEDA
 from utils.dataset_preparation import prepare_dataset
 from utils.preprocessing import preprocess_images
-from utils.augmentations import augment_images
-from utils.feature_extraction import feature_extraction_pipeline
-from utils.classifier import feature_classification_pipeline
+from utils.augment_images import augment_images
+from utils.feature_extractor import feature_extraction_pipeline
+from utils.classifier import train_and_evaluate
 
 
 def main():
@@ -27,24 +27,24 @@ def main():
     AUGMENTATION_STEPS = config_data.dataset_preparation.augmentation_steps  # "translation"  # Options: 'translation', 'rotation', 'gaussian_noise'
     HOW_MANY_AUGMENTATIONS = config_data.dataset_preparation.how_many_augmentations  # 4
     TRAIN_SET_RATIO = config_data.dataset_preparation.train_set_ratio  # 0.24: 50 - train, 12 - test
-    NORMALIZE_DEFAULT = config_data.dataset_preparation.normalize_default  # True False
+    IS_NORMALIZE_DEFAULT = config_data.dataset_preparation.normalize_default  # True False
     NORMALIZATION_METHOD = config_data.dataset_preparation.normalization_method  # min-max
 
-    NORMALIZE_WHEN_PREPROCESS = config_data.preprocessing_params.normalize_when_preprocess
-    DO_BRAIN_EXTRACTION = config_data.preprocessing_params.do_brain_exaction
-    DO_CROP = config_data.preprocessing_params.do_crop
-    DO_SMOOTH = config_data.preprocessing_params.do_smooth
-    DO_RE_NORMALIZE_AFTER_SMOOTH = config_data.preprocessing_params.do_re_normalize_after_smooth
-    PREPROCESS_TEST_SET = config_data.preprocessing_params.preprocess_test_set
+    IS_NORMALIZE_WHEN_PREPROCESS = config_data.preprocessing_params.normalize_when_preprocess
+    IS_BRAIN_EXTRACTION = config_data.preprocessing_params.do_brain_exaction
+    IS_CROP = config_data.preprocessing_params.do_crop
+    IS_SMOOTH = config_data.preprocessing_params.do_smooth
+    IS_RE_NORMALIZE_AFTER_SMOOTH = config_data.preprocessing_params.do_re_normalize_after_smooth
+    IS_PREPROCESS_TEST_SET = config_data.preprocessing_params.preprocess_test_set
 
[…]

### 2025-02-14 src/utils/classifier.py

@@ -10,13 +10,13 @@ logging.basicConfig(level=logging.INFO,
                     format="%(asctime)s - %(levelname)s - %(message)s")
 
 # Define paths
-FEATURES_DIR = "path/to/save/features"
-MODEL_PATH = "path/to/save/model.pkl"
+#FEATURES_DIR = "path/to/save/features"
+#MODEL_PATH = "path/to/save/model.pkl"
 
 
 def load_features(
-    features_dir: str = FEATURES_DIR
-) -> tuple[np.ndarray, np.ndarray, np.ndarray, np.ndarray]:
+    features_dir: str
+    ) -> tuple[np.ndarray, np.ndarray, np.ndarray, np.ndarray]:
     """
     Loads extracted features and labels from disk.
 
@@ -30,11 +30,9 @@ def load_features(
         FileNotFoundError: If any of the required feature files are missing.
     """
     try:
-        train_features = np.load(
-            os.path.join(features_dir, "train_features.npy"))
+        train_features = np.load(os.path.join(features_dir, "train_features.npy"))
         train_labels = np.load(os.path.join(features_dir, "train_labels.npy"))
-        test_features = np.load(os.path.join(features_dir,
-                                             "test_features.npy"))
+        test_features = np.load(os.path.join(features_dir, "test_features.npy"))
         test_labels = np.load(os.path.join(features_dir, "test_labels.npy"))
     except FileNotFoundError as e:
         logging.error(f"Feature file not found: {e}")
@@ -43,8 +41,12 @@ def load_features(
     return train_features, train_labels, test_features, test_labels
 
 
-def train_and_evaluate(model_path: str = MODEL_PATH,
-                       features_dir: str = FEATURES_DIR) -> None:
+def train_and_evaluate(extracted_features_dir: str,
+                       dir_to_save_clf: str,
+                       results_output_dir: str,
+                       clf_kernel,
+                       clf_c_value,
+                       clf_gamma_value) -> None:
     """
     Trains an SVM classifier on extracted features and evaluates it on a test set.
 
@@ -56,16 +58,19 @@ def train_and_evaluate(model_path: str = MODEL_PATH,
         None
     """
     logging.info("Loading extracted features...")
-    X_train, y_train, X_test, y_test = load_features(features_dir)
+    X_train, y_train, X_test, y_test = load_features(extracted_features_dir)
 
     # Train SVM classifier
     logging.info("Training SVM classifier...")
-    classifier = SVMClassifier()
[…]

### 2025-02-14 src/utils/eda.py
add doctrings to eda.py

@@ -1,79 +1,160 @@
+import os
 import pandas as pd
-import numpy as np
 import matplotlib.pyplot as plt
-import seaborn as sns
 import torch
+from typing import Optional, Dict, List, Any
 
 from utils.preprocess_validation import calculate_metrics
 
-class SchizophreniaEDA:
-    def __init__(self, path_to_clinical_data = None, path_to_raw_images = None):
-        """Initializes the EDA class with the dataset."""
-        self.df = pd.read_csv(path_to_clinical_data)
-        self.schiz_df = self._add_schiz_column()
-        self._create_age_groups()
 
-        self.path_to_raw_images = path_to_raw_images
-    
-    def _add_schiz_column(self):
-        """Adds a binary schizophrenia column."""
+class SchizophreniaEDA:
+    """Performs exploratory data analysis (EDA) on the clinical data and calculates image metrics"""
+
+    def __init__(self,
+                 path_to_clinical_data: str,
+                 path_to_raw_images: Optional[str] = None) -> None:
+        """
+        Initializes the SchizophreniaEDA class with paths to clinical and raw imaging data.
+
+        Parameters:
+            path_to_clinical_data (str): Path to the CSV file containing clinical data.
+            path_to_raw_images (Optional[str]): Directory containing raw MRI images in `.pt` format.
+            
+        Returns:
+            None.
+        """
+        self.df: pd.DataFrame = pd.read_csv(path_to_clinical_data)
+        self.schiz_df: pd.DataFrame = self._add_schiz_column()
+        self._create_age_groups()
+        self.path_to_raw_images: Optional[str] = path_to_raw_images
+
+    def _add_schiz_column(self) -> pd.DataFrame:
+        """
+        Adds a binary schizophrenia column to the clinical dataset.
+
+        Returns:
+            pd.DataFrame: Dataframe with an additional `schiz` column.
+        """
         df = self.df.copy()
-        df["schiz"] = df["dx"].apply(lambda x: "Schizophrenia" if "Schizophrenia" in x else "No Schizophrenia")
+        df["schiz"] = df["dx"].apply(
+            lambda x: "Schizophrenia"
+            if "Schizophrenia" in x else "No Schizophrenia")
         return df
-    
-    def _create_age_groups(self):
-        """Creates age group bins."""
+
+    def _create_age_groups(self) -> None:
+        """categorical age groups for better visualization"""
         bins = [0, 20, 30, 40, 50, 60, float('inf')]
[…]

### 2025-02-14 src/utils/feature_extractor.py
update feature_extractor

@@ -65,21 +65,24 @@ def feature_extraction_pipeline(train_set_dir,
                              collate_fn=collate_fn)
 
     device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
-    model = FeatureExtractor(
-        base_model_name=feature_extractor_model,
-        weights=weights,
-        input_channels=input_channels).to(device)
+    model = FeatureExtractor(base_model_name=feature_extractor_model,
+                             weights=weights,
+                             input_channels=input_channels).to(device)
 
     # Extract features
     print("Extracting training features...")
     train_features, train_labels = extract_features(train_loader, model,
                                                     device)
-    np.save(os.path.join(extracted_features_dir, "train_features.npy"), train_features)
-    np.save(os.path.join(extracted_features_dir, "train_labels.npy"), train_labels)
+    np.save(os.path.join(extracted_features_dir, "train_features.npy"),
+            train_features)
+    np.save(os.path.join(extracted_features_dir, "train_labels.npy"),
+            train_labels)
 
     print("Extracting test features...")
     test_features, test_labels = extract_features(test_loader, model, device)
-    np.save(os.path.join(extracted_features_dir, "test_features.npy"), test_features)
-    np.save(os.path.join(extracted_features_dir, "test_labels.npy"), test_labels)
+    np.save(os.path.join(extracted_features_dir, "test_features.npy"),
+            test_features)
+    np.save(os.path.join(extracted_features_dir, "test_labels.npy"),
+            test_labels)
 
     print("Feature extraction completed.")

### 2025-02-14 src/main.py
updated config, workflow and minors

@@ -1,3 +1,4 @@
+import os
 from logger import logger
 from omegaconf import OmegaConf
 from utils.eda import SchizophreniaEDA
@@ -6,45 +7,39 @@ from utils.preprocessing import preprocess_images
 from utils.augment_images import augment_images
 from utils.feature_extractor import feature_extraction_pipeline
 from utils.classifier import train_and_evaluate
+from utils.plot_svm_metrics import SVMVisualizer
 
 
 def main():
     logger.info("The program started...")
 
     # Load configuration from a config file
-    config_data = OmegaConf.load('config.yaml')
+    config_data = OmegaConf.load('config/config.yaml')
 
-    # Accessing parameters # TODO: USE CAPITALS FOR CONSTANTS
+    # Accessing parameters
     CLINICAL_DATA_DIR = config_data.file_paths.clinical_data_dir
     RAW_PT_DATA_DIR = config_data.file_paths.raw_pt_data_dir
     RAW_NII_DATA_DIR = config_data.file_paths.raw_nii_data_dir
-    PREPROCESSED_DATA_DIR = config_data.file_paths.preprocessed_dir
-    AUGMENTED_DATA_DIR = config_data.file_paths.augmented_dir
-    TRAIN_SET_DIR = config_data.file_paths.train_set_dir
-    TEST_SET_DIR = config_data.file_paths.test_set_dir
 
-    PREPROCESSING_STEPS = config_data.dataset_preparation.preprocessing_steps  # ["resample", "normalize", "extract_brain"]
-    AUGMENTATION_STEPS = config_data.dataset_preparation.augmentation_steps  # "translation"  # Options: 'translation', 'rotation', 'gaussian_noise'
-    HOW_MANY_AUGMENTATIONS = config_data.dataset_preparation.how_many_augmentations  # 4
+    #PREPROCESSING_STEPS = config_data.dataset_preparation.preprocessing_steps  # ["resample", "normalize", "extract_brain"]
+    #AUGMENTATION_STEPS = config_data.dataset_preparation.augmentation_steps  # "translation"  # Options: 'translation', 'rotation', 'gaussian_noise'
+    TRAIN_SET_DIR = config_data.dataset_preparation.train_set_dir
+    TEST_SET_DIR = config_data.dataset_preparation.test_set_dir
     TRAIN_SET_RATIO = config_data.dataset_preparation.train_set_ratio  # 0.24: 50 - train, 12 - test
[…]

### 2025-02-14 src/utils/classifier.py

@@ -1,17 +1,9 @@
 import os
 import json
 import numpy as np
-import logging
-from sklearn.metrics import accuracy_score, classification_report, roc_auc_score, confusion_matrix
+from logger import logger
 from src.models.svm import SVMClassifier
-
-# Configure logging
-logging.basicConfig(level=logging.INFO,
-                    format="%(asctime)s - %(levelname)s - %(message)s")
-
-# Define paths
-#FEATURES_DIR = "path/to/save/features"
-#MODEL_PATH = "path/to/save/model.pkl"
+from sklearn.metrics import accuracy_score, classification_report, roc_auc_score, confusion_matrix
 
 
 def load_features(
@@ -35,7 +27,7 @@ def load_features(
         test_features = np.load(os.path.join(features_dir, "test_features.npy"))
         test_labels = np.load(os.path.join(features_dir, "test_labels.npy"))
     except FileNotFoundError as e:
-        logging.error(f"Feature file not found: {e}")
+        logger.error(f"Feature file not found: {e}")
         raise
 
     return train_features, train_labels, test_features, test_labels
@@ -57,11 +49,11 @@ def train_and_evaluate(extracted_features_dir: str,
     Returns:
         None
     """
-    logging.info("Loading extracted features...")
+    logger.info("Loading extracted features...")
     X_train, y_train, X_test, y_test = load_features(extracted_features_dir)
 
     # Train SVM classifier
-    logging.info("Training SVM classifier...")
+    logger.info("Training SVM classifier...")
     classifier = SVMClassifier(kernel=clf_kernel,
                                C=clf_c_value,
                                gamma=clf_gamma_value,
@@ -70,11 +62,11 @@ def train_and_evaluate(extracted_features_dir: str,
 
     # Save trained model
     classifier.save_model(dir_to_save_clf)
-    logging.info(f"Model saved to {dir_to_save_clf}")
+    logger.info(f"Model saved to {dir_to_save_clf}")
 
     # Make predictions and get decision function for ROC AUC
     test_preds = classifier.predict(X_test)
-    test_scores = classifier.decision_function(X_test)
+    test_scores = classifier.predict_proba(X_test)
 
     # Compute evaluation metrics
     acc = accuracy_score(y_test, test_preds)
@@ -87,11 +79,14 @@ def train_and_evaluate(extracted_features_dir: str,
         "accuracy": acc,
         "roc_auc": auc,
         "classification_report": report,
[…]

### 2025-02-14 src/utils/dataset_preparation.py

@@ -5,7 +5,7 @@ import torch
 from src.utils.preprocess import load_nii, resample_image, normalize_data
 
 
-def prepare_dataset(train_ratio: float = 0.8,
+def prepare_dataset(train_ratio: float = 0.76,
                     raw_pt_dir: str = "data/raw_pt",
                     raw_nii_dir: str = "data/raw_nii",
                     train_set_output_dir: str = "train_set",

### 2025-02-14 src/utils/plot_svm_metrics.py

@@ -1,29 +1,62 @@
+import os
+import json
+import numpy as np
 import matplotlib.pyplot as plt
 import seaborn as sns
 from sklearn.metrics import confusion_matrix, roc_curve, auc
 
+
 class SVMVisualizer:
-    def __init__(self):
-        """Initializes the visualization tools for SVM classification results."""
-        pass
-    
-    def plot_confusion_matrix(self, true_labels, predicted_labels, class_names=None):
+    def __init__(self, results_json_path):
+        """
+        Initializes the visualization class and loads classification results.
+
+        Parameters:
+            results_json_path (str): Path to the JSON file containing classification results.
+        """
+        self.results_json_path = results_json_path
+        self.results = self._load_results()
+
+    def _load_results(self):
+        """Loads classification results from the JSON file."""
+        if not os.path.exists(self.results_json_path):
+            raise FileNotFoundError(f"Results file not found: {self.results_json_path}")
+
+        with open(self.results_json_path, "r") as f:
+            return json.load(f)
+
+    def plot_confusion_matrix(self):
         """Plots the confusion matrix."""
-        cm = confusion_matrix(true_labels, predicted_labels)
+        y_test = np.array(self.results["y_test"])
+        test_preds = np.array(self.results["test_preds"])
+        conf_matrix = np.array(self.results["confusion_matrix"])
+
+        # Get unique class names if labels are more than just 0 and 1
+        unique_classes = sorted(set(y_test))
+
         plt.figure(figsize=(6, 5))
-        sns.heatmap(cm, annot=True, fmt='d', cmap='Blues', xticklabels=class_names, yticklabels=class_names)
+        sns.heatmap(conf_matrix, annot=True, fmt='d', cmap='Blues', 
+                    xticklabels=unique_classes, yticklabels=unique_classes)
         plt.xlabel('Predicted Labels')
         plt.ylabel('True Labels')
         plt.title('Confusion Matrix')
         plt.show()
-    
-    def plot_roc_curve(self, true_labels, test_scores):
+
+    def plot_roc_curve(self):
         """Plots the ROC curve."""
-        fpr, tpr, _ = roc_curve(true_labels, test_scores[:, 1])
-        roc_auc = auc(fpr, tpr)
+        y_test = np.array(self.results["y_test"])
[…]

### 2025-02-14 src/main.py
updated workflow explanations

@@ -17,9 +17,11 @@ def main():
     config_data = OmegaConf.load('config/config.yaml')
 
     # Accessing parameters
-    CLINICAL_DATA_DIR = config_data.file_paths.clinical_data_dir
-    RAW_PT_DATA_DIR = config_data.file_paths.raw_pt_data_dir
-    RAW_NII_DATA_DIR = config_data.file_paths.raw_nii_data_dir
+    RAW_PT_DATA_DIR = config_data.base.raw_pt_data_dir
+    RAW_NII_DATA_DIR = config_data.base.raw_nii_data_dir
+
+    CLINICAL_DATA_DIR = config_data.eda.clinical_data_dir
+    IS_PERFORM_EDA = config_data.eda.is_perform_eda
 
     #PREPROCESSING_STEPS = config_data.dataset_preparation.preprocessing_steps  # ["resample", "normalize", "extract_brain"]
     #AUGMENTATION_STEPS = config_data.dataset_preparation.augmentation_steps  # "translation"  # Options: 'translation', 'rotation', 'gaussian_noise'
@@ -72,24 +74,30 @@ def main():
 
     # Step 0: Explanatory data analysis
     """
-    Performs an explanatory data analysis on the raw images and corresponding clinical data.
+    Performs an explanatory data analysis on the raw images (data/raw_pt)
+    and corresponding clinical data (data/clinical_data.csv).
     This step is optional and can be skipped if not necessary.
     """
-    eda = SchizophreniaEDA(CLINICAL_DATA_DIR, RAW_PT_DATA_DIR)
-    eda.plot_age_distribution()
-    eda.plot_gender_distribution()
-    eda.get_age_statistics()
-    eda.save_raw_images_metrics()
-    logger.info("EDA completed.")
+    if IS_PERFORM_EDA:
+        eda = SchizophreniaEDA(CLINICAL_DATA_DIR, RAW_PT_DATA_DIR)
+        eda.plot_age_distribution()
+        eda.plot_gender_distribution()
+        eda.get_age_statistics()
+        eda.save_raw_images_metrics()
+        logger.info("EDA completed.")
+    else:
+        logger.info("EDA was skipped.")    
 
     # Step 1: Select the data (train, test)
     """
-    Randomly divides (ratio is user defined) images located in /data/raw,
-    Creates 2 folders: train_set, test_set,
+    Randomly divides (ratio is user defined) images located in /data/raw_nii,
[…]

### 2025-02-14 src/utils/feature_extractor.py

@@ -7,8 +7,6 @@ from torch.utils.data import DataLoader
 from src.utils.data_loader import MRIDataset
 from src.models.cnn import FeatureExtractor
 
-# TODO: ADD CORRECT TRAIN_SET_DIR, TEST_SET_DIR, LABELS_FILE, FEATURE_EXTRACTOR_MODEL, BATCH_SIZE, FEATURES_DIR
-
 
 # Collate function for DataLoader (handles empty batches)
 def collate_fn(batch):

### 2025-02-14 src/main.py
completed workflow

@@ -61,8 +61,9 @@ def main():
     BASE_MODEL_WEIGHTS = config_data.resnet_params.base_model_weights
     INPUT_CHANNELS = config_data.resnet_params.input_channels
 
-    LABELS_DIR = config_data.feature_extraction_params.labels_dir
+    LABELS_DIR = config_data.eda.clinical_data_dir
     EXTRACTED_FEATURES_DIR = config_data.feature_extraction_params.extracted_features_dir
+    TARGET_SHAPE = config_data.feature_extraction_params.target_shape
     BATCH_SIZE = config_data.feature_extraction_params.batch_size
 
     SAVE_CLF_DIR = config_data.svc_params.save_clf_dir
@@ -125,7 +126,10 @@ def main():
     may also be resampled and normalized for consistency (user-defined option),
     Saves preprocessed test images in data/raw_nii/test_set/preprocessed.
     """
-    preprocess_images(normalize=IS_NORMALIZE_WHEN_PREPROCESS,
+    preprocess_images(raw_nii_dir=RAW_NII_DATA_DIR,
+                      train_set_dir=TRAIN_SET_DIR,
+                      test_set_dir=TEST_SET_DIR,
+                      normalize=IS_NORMALIZE_WHEN_PREPROCESS,
                       norm_metod=NORMALIZATION_METHOD,
                       min_max_min_val=MIN_VAL,
                       min_max_max_val=MAX_VAL,
@@ -165,8 +169,10 @@ def main():
             }),
         ],
         num_augmentations=HOW_MANY_AUGMENTATIONS,
-        input_dir=PREPROCESSED_DATA_DIR,
-        output_dir=AUGMENTED_DATA_DIR,
+        raw_nii_dir=RAW_NII_DATA_DIR,
+        train_set_dir=TRAIN_SET_DIR,
+        preprocessed_train_set_dir=PREPROCESSED_DATA_DIR,
+        output_dir=AUGMENTED_DATA_DIR
     )
     logger.info("Augmentation completed.")
 
@@ -177,10 +183,14 @@ def main():
     data/raw_nii/train_set/preprocessed and data/raw_nii/train_set/preprocessed/augmented
     Saves the extracted features in data/raw_nii/train_set/extracted_features.
     """
-    feature_extraction_pipeline(train_set_dir=TRAIN_SET_DIR,
+    feature_extraction_pipeline(raw_nii_dir=RAW_NII_DATA_DIR,
[…]

### 2025-02-14 src/utils/augment_images.py

@@ -4,10 +4,12 @@ import random
 from src.utils.augmentations import apply_translation, apply_rotation, apply_gaussian_noise
 
 def augment_images(
-        augmentations,
-        num_augmentations,
-        input_dir="data/raw_nii/train_set/preprocessed",
-        output_dir="data/raw_nii/train_set/preprocessed/augmented") -> None:
+        augmentations: list[tuple],
+        num_augmentations: int = 3,
+        raw_nii_dir: str = "data/raw_nii",
+        train_set_dir: str = "train_set",
+        preprocessed_train_set_dir: str = "preprocessed",
+        output_dir="augmented") -> None:
     """
     Applies user-specified augmentations to images in input_dir and saves them to output_dir.
 
@@ -21,7 +23,8 @@ def augment_images(
     Returns:
         None
     """
-
+    input_dir = os.path.join(raw_nii_dir, train_set_dir, preprocessed_train_set_dir)
+    output_dir = os.path.join(input_dir, output_dir)
     os.makedirs(output_dir, exist_ok=True)
 
     # Get images from folder and load them

### 2025-02-14 src/utils/classifier.py

@@ -33,12 +33,12 @@ def load_features(
     return train_features, train_labels, test_features, test_labels
 
 
-def train_and_evaluate(extracted_features_dir: str,
-                       dir_to_save_clf: str,
-                       results_output_dir: str,
-                       clf_kernel,
-                       clf_c_value,
-                       clf_gamma_value) -> None:
+def train_and_evaluate(extracted_features_dir: str = "data/extracted_features",
+                       dir_to_save_clf: str = "src/models/svm_classifier.pkl",
+                       results_output_dir: str = "src/models/results/classification_results.json",
+                       clf_kernel: str = "rbf",
+                       clf_c_value: int = 100,
+                       clf_gamma_value: float = 0.0001) -> None:
     """
     Trains an SVM classifier on extracted features and evaluates it on a test set.
 
@@ -85,8 +85,7 @@ def train_and_evaluate(extracted_features_dir: str,
         "test_scores": test_scores.tolist()  # For ROC AUC visualization
     }
 
-    results_path = os.path.join(extracted_features_dir, results_output_dir+'.json') # "classification_results.json"
-    with open(results_path, "w") as f:
+    with open(results_output_dir, "w") as f:
         json.dump(results, f, indent=4)
 
-    logger.info(f"Classification results saved to {results_path}")
+    logger.info(f"Classification results saved to {results_output_dir}")

### 2025-02-14 src/utils/feature_extractor.py

@@ -1,11 +1,23 @@
 import os
 import torch
+import glob
 import numpy as np
 import pandas as pd
 import tqdm
 from torch.utils.data import DataLoader
 from src.utils.data_loader import MRIDataset
 from src.models.cnn import FeatureExtractor
+
+
+# Function to get image paths
+def get_image_paths(folder_path):
+    """Gets all .pt file paths from the specified folder."""
+    image_paths = [
+        os.path.join(folder_path, file_name)
+        for file_name in os.listdir(folder_path)
+        if file_name.endswith('.pt')
+    ]
+    return image_paths
 
 
 # Collate function for DataLoader (handles empty batches)
@@ -39,19 +51,40 @@ def extract_features(loader, model, device):
 
 
 # Main feature extraction pipeline
-def feature_extraction_pipeline(train_set_dir,
-                                test_set_dir,
-                                labels_dir,
-                                extracted_features_dir,
-                                batch_size,
-                                feature_extractor_model,
-                                weights,
+def feature_extraction_pipeline(raw_nii_dir: str = "data/raw_nii",
+                                train_set_dir: str = "train_set",
+                                test_set_dir: str = "test_set",
+                                preprocessed_set_dir: str = "preprocessed",
+                                augmented_train_set_dir: str = "augmented",
+                                labels_dir: str = "data/clinical_data.csv",
+                                extracted_features_dir: str = "data/extracted_features",
+                                target_shape: tuple = (224, 224),
+                                batch_size: int = 32,
+                                feature_extractor_model: str = 'resnet18',
+                                base_model_weights = None,
                                 input_channels: int = 1) -> None:
     os.makedirs(extracted_features_dir, exist_ok=True)
 
+    '''preprocessed_train_data_dir = os.path.join(raw_nii_dir, train_set_dir, preprocessed_set_dir)
+    augmented_train_data_dir = os.path.join(raw_nii_dir, train_set_dir, preprocessed_set_dir, augmented_train_set_dir)
+    
+    # Collect paths from both preprocessed and augmented directories
+    preprocessed_paths = get_image_paths(preprocessed_train_data_dir)
+    augmented_paths = get_image_paths(augmented_train_data_dir)
+
+    # Combine both lists
+    all_train_set_paths = preprocessed_paths + augmented_paths'''
+
[…]

### 2025-02-14 src/utils/preprocessing.py

@@ -3,6 +3,9 @@ import torch
 from src.utils.preprocess import normalize_data, extract_brain, crop_to_largest_bounding_box, apply_gaussian_smoothing
 
 def preprocess_images(
+        raw_nii_dir: str = "data/raw_nii",
+        train_set_dir: str = "train_set",
+        test_set_dir: str = "test_set",
         normalize: bool = False,
         norm_method: str = "min-max",
         min_max_min_val: float = 0,
@@ -23,9 +26,9 @@ def preprocess_images(
         ) -> None:
     """
     """
-    for dataset in ["train_set", "test_set"] if preprocess_test_set else ["train_set"]:
-        input_dir = f"data/raw_nii/{dataset}"
-        output_dir = output_dir #f"{input_dir}/preprocessed"
+    for dataset in [train_set_dir, test_set_dir] if preprocess_test_set else [train_set_dir]:
+        input_dir = raw_nii_dir + '/' + dataset #f"data/raw_nii/{dataset}"
+        output_dir = input_dir + '/' + output_dir #f"{input_dir}/preprocessed"
         os.makedirs(output_dir, exist_ok=True)
         
         for img in os.listdir(input_dir):

### 2025-02-14 src/main.py
minor changes

@@ -1,4 +1,3 @@
-import os
 from logger import logger
 from omegaconf import OmegaConf
 from utils.eda import SchizophreniaEDA

### 2025-02-15 tests/test_data_loader.py
new tests

@@ -1,50 +1,68 @@
-import pytest
+import unittest
 import torch
-import os
 import pandas as pd
-import glob
-from unittest.mock import patch, MagicMock
-from utils.def_preprocess import load_nii
-from utils.def_data_loader import MRIDataset 
+import tempfile
+from pathlib import Path
+from utils.data_loader import MRIDataset  # Ensure this is correctly imported
 
-def mock_load_nii(file_path):
-    """Mock function to simulate loading a NIfTI image."""
-    return load_nii(torch.rand((1, 128, 128, 128)))  # Simulating a 3D tensor
+class TestMRIDataset(unittest.TestCase):
+    def setUp(self):
+        # Create temporary directory for fake MRI tensor files
+        self.temp_dir = tempfile.TemporaryDirectory()
+        self.mri_files = []
 
-def mock_preprocess_image(image):
-    """Mock function to simulate preprocessing an image."""
-    return image  # Assume it returns the same tensor
+        # Generate synthetic MRI tensors and save them
+        for i in range(3):  # 3 subjects
+            subject_id = f"sub-A10{i}"  # Matches the regex
+            file_path = Path(self.temp_dir.name) / f"{subject_id}.pt"
+            tensor_data = torch.rand((10, 128, 128))  # 10 slices, 128x128 each
+            torch.save(tensor_data, file_path)
+            self.mri_files.append(str(file_path))
 
-@patch("src.utils.data_loader", side_effect=mock_load_nii)
-@patch("src.utils.preprocess", side_effect=mock_preprocess_image)
-@patch("pandas.read_csv")
-@patch("glob.glob")
-def test_mri_dataset(mock_glob, mock_read_csv, mock_preprocess, mock_load_nii):
-    """Test the MRIDataset class."""
-    # Mock the CSV file
-    mock_read_csv.return_value = pd.DataFrame({
-        "participant_id": ["sub-A000001", "sub-A000002"],
-        "dx_encoded": [0, 1]
-    }).set_index("participant_id")
+        # Create a fake labels DataFrame
+        self.labels_df = pd.DataFrame({
+            "participant_id": ["A100", "A101", "A102"],
+            "dx_encoded": [0, 1, 1]  # Example labels
+        })
 
-    # Mock the file paths found by glob
-    mock_glob.return_value = [
-        "/fake_path/schizconnect_COBRE_images_22613/COBRE/sub-A000001/ses-01/anat/image1.nii.gz",
[…]

### 2025-02-15 src/main.py
configurable augmentations

@@ -51,8 +51,11 @@ def main():
 
     AUGMENTED_DATA_DIR = config_data.augmentation_params.augmented_data_dir
     HOW_MANY_AUGMENTATIONS = config_data.augmentation_params.how_many_augmentations  # 4
+    IS_TRANSLATION = config_data.augmentation_params.is_translation
     TRANSLATION_SHIFT = config_data.augmentation_params.translation_shift
+    IS_ROTATION = config_data.augmentation_params.is_rotation
     ROTATION_ANGLE = config_data.augmentation_params.rotation_angle
+    IS_GAUSSIAN_NOISE = config_data.augmentation_params.is_gaussian_noise
     GAUSSIAN_NOISE_MEAN = config_data.augmentation_params.gaussian_noise_mean
     GAUSSIAN_NOISE_STD = config_data.augmentation_params.gaussian_noise_std
 
@@ -71,6 +74,7 @@ def main():
     C = config_data.svc_params.C
     GAMMA = config_data.svc_params.gamma
 
+    FIGS_OUTPUT_DIR = config_data.plot_params.figs_output_dir
 
     # Step 0: Explanatory data analysis
     """
@@ -87,6 +91,7 @@ def main():
         logger.info("EDA completed.")
     else:
         logger.info("EDA was skipped.")    
+
 
     # Step 1: Select the data (train, test)
     """
@@ -110,6 +115,7 @@ def main():
                     min_max_min_val=MIN_VAL,
                     min_max_max_val=MAX_VAL)
     logger.info("Dataset preparation completed.")
+
 
     # Step 2: Preprocessing
     """
@@ -148,25 +154,27 @@ def main():
                       )
     logger.info("Preprocessing completed.")
 
+
     # Step 3: Augmentation
     """
     Applies specified augmentation type (translation, rotation, gaussian_noise, or combined)
     to the preprocessed train set images located in data/raw_nii/train_set/preprocessed,
     Saves augmented images in data/raw_nii/train_set/preprocessed/augmented.
     """
+    # Construct augmentation list based on config
+    augmentations = []
+
+    if IS_TRANSLATION:
+        augmentations.append(("translation", {"shift": TRANSLATION_SHIFT}))
+
+    if IS_ROTATION:
+        augmentations.append(("rotation", {"angle": ROTATION_ANGLE}))
+
+    if IS_GAUSSIAN_NOISE:
[…]

### 2025-02-15 src/utils/plot_svm_metrics.py

@@ -7,7 +7,7 @@ from sklearn.metrics import confusion_matrix, roc_curve, auc
 
 
 class SVMVisualizer:
-    def __init__(self, results_json_path):
+    def __init__(self, results_json_path, save_figs_path:str = "paper/figs"):
         """
         Initializes the visualization class and loads classification results.
 
@@ -16,6 +16,7 @@ class SVMVisualizer:
         """
         self.results_json_path = results_json_path
         self.results = self._load_results()
+        self.save_figs_path = save_figs_path
 
     def _load_results(self):
         """Loads classification results from the JSON file."""
@@ -34,13 +35,14 @@ class SVMVisualizer:
         # Get unique class names if labels are more than just 0 and 1
         unique_classes = sorted(set(y_test))
 
-        plt.figure(figsize=(6, 5))
+        fig = plt.figure(figsize=(6, 5))
         sns.heatmap(conf_matrix, annot=True, fmt='d', cmap='Blues', 
                     xticklabels=unique_classes, yticklabels=unique_classes)
         plt.xlabel('Predicted Labels')
         plt.ylabel('True Labels')
         plt.title('Confusion Matrix')
-        plt.show()
+        fig.savefig(self.save_figs_path+"cm"+".png")
+        #plt.show()
 
     def plot_roc_curve(self):
         """Plots the ROC curve."""
@@ -55,11 +57,12 @@ class SVMVisualizer:
 
         roc_auc = auc(fpr, tpr)
 
-        plt.figure(figsize=(6, 5))
+        fig = plt.figure(figsize=(6, 5))
         plt.plot(fpr, tpr, color='blue', label=f'ROC curve (AUC = {roc_auc:.2f})')
         plt.plot([0, 1], [0, 1], color='gray', linestyle='--')
         plt.xlabel('False Positive Rate')
         plt.ylabel('True Positive Rate')
         plt.title('Receiver Operating Characteristic (ROC) Curve')
         plt.legend()
-        plt.show()
+        fig.savefig(self.save_figs_path+"roc"+".png")
+        #plt.show()

### 2025-02-15 src/main.py
minor improvements

@@ -134,22 +134,22 @@ def main():
     preprocess_images(raw_nii_dir=RAW_NII_DATA_DIR,
                       train_set_dir=TRAIN_SET_DIR,
                       test_set_dir=TEST_SET_DIR,
-                      normalize=IS_NORMALIZE_WHEN_PREPROCESS,
+                      is_normalize=IS_NORMALIZE_WHEN_PREPROCESS,
                       norm_metod=NORMALIZATION_METHOD,
                       min_max_min_val=MIN_VAL,
                       min_max_max_val=MAX_VAL,
-                      brain_extraction=IS_BRAIN_EXTRACTION,
+                      is_brain_extraction=IS_BRAIN_EXTRACTION,
                       brain_extraction_modality=MODALITY,
                       brain_extraction_verbose=VERBOSE,
-                      crop=IS_CROP,
-                      smooth=IS_SMOOTH,
+                      is_crop=IS_CROP,
+                      is_smooth=IS_SMOOTH,
                       smooth_sigma=SIGMA,
                       smooth_order=ORDER,
                       smooth_mode=MODE,
                       smooth_cval=CVAL,
                       smooth_truncate=TRUNCATE,
-                      re_normalize_after_smooth=IS_RE_NORMALIZE_AFTER_SMOOTH,
-                      preprocess_test_set=IS_PREPROCESS_TEST_SET,
+                      is_re_normalize_after_smooth=IS_RE_NORMALIZE_AFTER_SMOOTH,
+                      is_preprocess_test_set=IS_PREPROCESS_TEST_SET,
                       output_dir=PREPROCESSED_DATA_DIR,
                       )
     logger.info("Preprocessing completed.")

### 2025-02-15 src/utils/preprocessing.py

@@ -1,67 +1,116 @@
 import os
 import torch
-from src.utils.preprocess import normalize_data, extract_brain, crop_to_largest_bounding_box, apply_gaussian_smoothing
+from src.utils.preprocess import (
+    normalize_data,
+    extract_brain,
+    crop_to_largest_bounding_box,
+    apply_gaussian_smoothing,
+)
 
 def preprocess_images(
-        raw_nii_dir: str = "data/raw_nii",
-        train_set_dir: str = "train_set",
-        test_set_dir: str = "test_set",
-        normalize: bool = False,
-        norm_method: str = "min-max",
-        min_max_min_val: float = 0,
-        min_max_max_val: float = 1,
-        brain_extraction: bool = True, 
-        brain_extraction_modality: str = "t1",
-        brain_extraction_verbose: bool = False,
-        crop: bool = False, 
-        smooth: bool = False,
-        smooth_sigma: float = 1.5,
-        smooth_oder: int = 2,
-        smooth_mode: str = "constant",
-        smooth_cval: int = 1,
-        smooth_truncate: float = 2, 
-        re_normalize_after_smooth: bool = False, 
-        preprocess_test_set: bool = False,
-        output_dir: str = "preprocessed",
-        ) -> None:
+    raw_nii_dir: str = "data/raw_nii",
+    train_set_dir: str = "train_set",
+    test_set_dir: str = "test_set",
+    is_normalize: bool = False,
+    norm_method: str = "min-max",
+    min_max_min_val: float = 0,
+    min_max_max_val: float = 1,
+    is_brain_extraction: bool = True, 
+    brain_extraction_modality: str = "t1",
+    brain_extraction_verbose: bool = False,
+    is_crop: bool = False, 
+    is_smooth: bool = False,
+    smooth_sigma: float = 1.5,
+    smooth_order: int = 2,
+    smooth_mode: str = "constant",
+    smooth_cval: int = 1,
+    smooth_truncate: float = 2, 
+    is_re_normalize_after_smooth: bool = False, 
+    is_preprocess_test_set: bool = False,
+    output_dir: str = "preprocessed",
+) -> None:
     """
+    Preprocesses MRI images based on user-defined steps.
+    
+    Supports: Normalization, Brain Extraction, Cropping, Smoothing, and Re-Normalization.
+
+    Parameters:
+        raw_nii_dir (str): Directory containing raw .nii images.
+        train_set_dir (str): Directory for training set.
+        test_set_dir (str): Directory for test set.
[…]

### 2025-02-15 src/utils/preprocess.py
all tests pass

@@ -325,7 +325,34 @@ def get_largest_brain_mask_slice(mask: np.ndarray) -> tuple[np.ndarray, int]:
 
     return binary_mask, largest_slice_index
 
-def crop_to_largest_bounding_box(data: np.ndarray, 
+def crop_to_largest_bounding_box(data: np.ndarray, mask: np.ndarray) -> np.ndarray:
+    """
+    Crop all slices to the bounding box of the largest brain area.
+    
+    Parameters:
+    - data: numpy array, 3D MRI data
+    - mask: numpy array, binary brain mask (3D)
+    
+    Returns:
+    - cropped_data: numpy array, cropped 3D MRI data
+    """
+    largest_slice_idx = np.argmax(np.sum(mask, axis=(0, 1)))  # Get slice with largest nonzero mask
+
+    largest_slice_mask = mask[:, :, largest_slice_idx]
+
+    coords = np.argwhere(largest_slice_mask > 0)
+    if coords.size == 0:  # If no brain region is found, return original data
+        print("Warning: No bounding box found. Returning original data.")
+        return data  # Or return np.zeros_like(data) if you prefer an empty image
+
+    x_min, y_min = coords.min(axis=0)
+    x_max, y_max = coords.max(axis=0) + 1  # Inclusive
+
+    cropped_slices = data[x_min:x_max, y_min:y_max, :]
+    return cropped_slices
+
+
+'''def crop_to_largest_bounding_box(data: np.ndarray, 
                                  processed_mask: np.ndarray = None, 
                                  largest_slice_idx: int = None,
                                  mask: np.ndarray = None
@@ -355,7 +382,7 @@ def crop_to_largest_bounding_box(data: np.ndarray,
     # Crop all slices using the bounding box dimensions
     cropped_slices = data[x_min:x_max, y_min:y_max, :]
 
-    return cropped_slices
+    return cropped_slices'''
 
 def apply_gaussian_smoothing(data, 
                              sigma=1.5, 

### 2025-02-15 src/utils/preprocess.py
tests pass

@@ -285,46 +285,6 @@ def extract_brain(data: np.ndarray,
     except Exception as e:
         raise RuntimeError(f"An unexpected error occurred: {str(e)}")  
 
-def get_largest_brain_mask_slice(mask: np.ndarray) -> tuple[np.ndarray, int]:
-    """
-    Determine the slices with the largest brain mask dimension,
-    using Otsu's thresholding for binarization.
-    
-    Parameters:
-    - mask: numpy array, brain mask (3D)
-    
-    Returns:
-    - largest_slice_index: int, index of the slice with the largest brain mask
-    """
-    # Apply Otsu's thresholding to binarize the mask
-    threshold = threshold_otsu(mask)
-    binary_mask = (mask > threshold).astype(np.uint8)
-
-    largest_area = 0
-    largest_slice_index = -1
-
-    # Iterate through each slice along the z-axis
-    for z in range(binary_mask.shape[2]):  # Loop through slices
-        slice_mask = binary_mask[:, :, z]
-
-        # Skip completely empty slices
-        if np.any(slice_mask > 0):
-            # Find the bounding box of the slice mask
-            coords = np.argwhere(slice_mask > 0)
-            x_min, y_min = coords.min(axis=0)
-            x_max, y_max = coords.max(axis=0) + 1  # Inclusive
-
-            height = x_max - x_min
-            width = y_max - y_min
-            area = height * width
-
-            # Update the largest slice index based on area
-            if area > largest_area:
-                largest_area = area
-                largest_slice_index = z
-
-    return binary_mask, largest_slice_index
-
 def crop_to_largest_bounding_box(data: np.ndarray, mask: np.ndarray) -> np.ndarray:
     """
     Crop all slices to the bounding box of the largest brain area.
@@ -350,39 +310,6 @@ def crop_to_largest_bounding_box(data: np.ndarray, mask: np.ndarray) -> np.ndarr
 
     cropped_slices = data[x_min:x_max, y_min:y_max, :]
     return cropped_slices
-
-
-'''def crop_to_largest_bounding_box(data: np.ndarray, 
-                                 processed_mask: np.ndarray = None, 
-                                 largest_slice_idx: int = None,
-                                 mask: np.ndarray = None
-                                 ) -> np.ndarray:
-    """
-    Crop all slices to the bounding box of the largest brain area.
-    
-    Parameters:
-    - data: numpy array, 3D MRI data
[…]

### 2025-02-15 src/main.py
minor updates

@@ -22,14 +22,14 @@ def main():
     CLINICAL_DATA_DIR = config_data.eda.clinical_data_dir
     IS_PERFORM_EDA = config_data.eda.is_perform_eda
 
-    #PREPROCESSING_STEPS = config_data.dataset_preparation.preprocessing_steps  # ["resample", "normalize", "extract_brain"]
-    #AUGMENTATION_STEPS = config_data.dataset_preparation.augmentation_steps  # "translation"  # Options: 'translation', 'rotation', 'gaussian_noise'
+    IS_PERFORM_DATASET_PREPARATION = config_data.dataset_preparation.is_perform_dataset_preparation
     TRAIN_SET_DIR = config_data.dataset_preparation.train_set_dir
     TEST_SET_DIR = config_data.dataset_preparation.test_set_dir
     TRAIN_SET_RATIO = config_data.dataset_preparation.train_set_ratio  # 0.24: 50 - train, 12 - test
     IS_NORMALIZE_WHEN_PREPARING = config_data.dataset_preparation.is_normalize_when_preparing  # True False
     NORMALIZATION_METHOD = config_data.dataset_preparation.normalization_method  # min-max
 
+    IS_PERFORM_PREPROCESSING = config_data.preprocessing_params.is_perform_preprocessing
     PREPROCESSED_DATA_DIR = config_data.preprocessing_params.preprocessed_data_dir
     IS_NORMALIZE_WHEN_PREPROCESS = config_data.preprocessing_params.is_normalize_when_preprocess
     IS_BRAIN_EXTRACTION = config_data.preprocessing_params.is_brain_exaction
@@ -49,6 +49,7 @@ def main():
     CVAL = config_data.preprocessing_params.cval
     TRUNCATE = config_data.preprocessing_params.truncate
 
+    IS_PERFORM_AUGMENTATION = config_data.augmentation_params.is_perform_augmentation
     AUGMENTED_DATA_DIR = config_data.augmentation_params.augmented_data_dir
     HOW_MANY_AUGMENTATIONS = config_data.augmentation_params.how_many_augmentations  # 4
     IS_TRANSLATION = config_data.augmentation_params.is_translation
@@ -59,6 +60,7 @@ def main():
     GAUSSIAN_NOISE_MEAN = config_data.augmentation_params.gaussian_noise_mean
     GAUSSIAN_NOISE_STD = config_data.augmentation_params.gaussian_noise_std
 
[…]

### 2025-02-15 src/main.py
logger support

@@ -206,7 +206,6 @@ def main():
                                     train_set_dir=TRAIN_SET_DIR,
                                     test_set_dir=TEST_SET_DIR,
                                     preprocessed_set_dir=PREPROCESSED_DATA_DIR,
-                                    augmented_train_set_dir=AUGMENTED_DATA_DIR,
                                     labels_dir=LABELS_DIR,
                                     extracted_features_dir=EXTRACTED_FEATURES_DIR,
                                     target_shape=TARGET_SHAPE,
@@ -253,7 +252,7 @@ def main():
         logger.info("Classifier metrics have been plotted.")
     else:
         logger.info("Plotting was skipped.")
-        
+
     logger.info("The program completed successfully.")
 
 if __name__ == "__main__":

### 2025-02-15 src/utils/dataset_preparation.py

@@ -2,6 +2,7 @@ import os
 import shutil
 import random
 import torch
+from logger import logger
 from src.utils.preprocess import load_nii, resample_image, normalize_data
 
 
@@ -77,6 +78,4 @@ def prepare_dataset(train_ratio: float = 0.76,
     process_and_save_nii(train_images, raw_nii_dir, train_nii_dir)
     process_and_save_nii(test_images, raw_nii_dir, test_nii_dir)
 
-    print(
-        f"Dataset prepared: {len(train_images)} images in train_set, {len(test_images)} images in test_set."
-    )
+    logger.info(f"Dataset prepared: {len(train_images)} images in train_set, {len(test_images)} images in test_set.")

### 2025-02-15 src/utils/eda.py

@@ -1,6 +1,7 @@
 import os
 import torch
 import pandas as pd
+from logger import logger
 import matplotlib.pyplot as plt
 from typing import Optional, Dict, List, Any
 from utils.preprocess_validation import calculate_metrics
@@ -134,7 +135,7 @@ class SchizophreniaEDA:
                 raw_metrics = calculate_metrics(None, tensor_numpy)
                 raw_metrics_all_images.append(raw_metrics)
             except Exception as e:
-                print(f"Error processing {file_path}: {e}")
+                logger.error(f"Error processing {file_path}: {e}")
 
         return raw_metrics_all_images
 
@@ -154,6 +155,6 @@ class SchizophreniaEDA:
         if raw_metrics_all_images:
             df_raw_metrics = pd.DataFrame(raw_metrics_all_images)
             df_raw_metrics.to_csv(save_path, index=False)
-            print(f"Raw image metrics saved to {save_path}")
+            logger.info(f"Raw image metrics saved to {save_path}")
         else:
-            print("No metrics computed. Check the input directory.")
+            logger.info("No metrics computed. Check the input directory.")

### 2025-02-15 src/utils/plot_svm_metrics.py

@@ -3,7 +3,7 @@ import json
 import numpy as np
 import matplotlib.pyplot as plt
 import seaborn as sns
-from sklearn.metrics import confusion_matrix, roc_curve, auc
+from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay, roc_curve, auc
 
 
 class SVMVisualizer:
@@ -33,14 +33,19 @@ class SVMVisualizer:
         conf_matrix = np.array(self.results["confusion_matrix"])
 
         # Get unique class names if labels are more than just 0 and 1
-        unique_classes = sorted(set(y_test))
+        #unique_classes = sorted(set(y_test))
 
         fig = plt.figure(figsize=(6, 5))
-        sns.heatmap(conf_matrix, annot=True, fmt='d', cmap='Blues', 
-                    xticklabels=unique_classes, yticklabels=unique_classes)
-        plt.xlabel('Predicted Labels')
-        plt.ylabel('True Labels')
-        plt.title('Confusion Matrix')
+        cm = confusion_matrix(y_test, test_preds)
+        disp = ConfusionMatrixDisplay(confusion_matrix=cm, display_labels=['Non-Schizophrenic', 'Schizophrenic'])
+        disp.plot(cmap='Blues')
+        plt.title("Confusion Matrix: Test Set")
+
+        #sns.heatmap(conf_matrix, annot=True, fmt='d', cmap='Blues', 
+        #            xticklabels=unique_classes, yticklabels=unique_classes)
+        #plt.xlabel('Predicted Labels')
+        #plt.ylabel('True Labels')
+        #plt.title('Confusion Matrix')
         fig.savefig(self.save_figs_path+"cm"+".png")
         #plt.show()
 

### 2025-02-15 src/utils/preprocess.py

@@ -4,6 +4,7 @@ import antspynet
 import numpy as np
 import nibabel as nib
 from typing import Union
+from logger import logger
 from scipy.ndimage import gaussian_filter
 from skimage.filters import threshold_otsu
 from nibabel.processing import resample_to_output 
@@ -302,7 +303,7 @@ def crop_to_largest_bounding_box(data: np.ndarray, mask: np.ndarray) -> np.ndarr
 
     coords = np.argwhere(largest_slice_mask > 0)
     if coords.size == 0:  # If no brain region is found, return original data
-        print("Warning: No bounding box found. Returning original data.")
+        logger.info("Warning: No bounding box found. Returning original data.")
         return data  # Or return np.zeros_like(data) if you prefer an empty image
 
     x_min, y_min = coords.min(axis=0)

### 2025-02-15 src/utils/preprocess_validation.py

@@ -2,6 +2,7 @@ from typing import Union
 import matplotlib.pyplot as plt
 import nibabel as nib
 import numpy as np
+from logger import logger
 from skimage.filters import threshold_otsu
 from skimage.metrics import mean_squared_error
 from skimage.metrics import structural_similarity as ssim
@@ -140,7 +141,7 @@ def generate_signal_mask(data: np.ndarray,
     # Filter non-zero values for Otsu's thresholding
     non_zero_data = data[data > 0]
     if non_zero_data.size == 0:
-        print(
+        logger.info(
             "Warning: No non-zero data found for thresholding. Returning a zero mask."
         )
         return np.zeros_like(data, dtype=np.uint8)
@@ -173,7 +174,7 @@ def calculate_snr_with_mask(data: np.ndarray,
 
     # Check if the mask has any valid signal region
     if np.sum(mask) == 0:
-        print(
+        logger.info(
             "Warning: The mask contains no signal region. Assigning SNR value of 0."
         )
         return 0.0

### 2025-02-15 src/paper/paper.tex
minor updates
update readme, update paper, update plot script

@@
         \centering
-        \caption{\label{tab:cobre_clinical_demographic}The subjects were selected to give a balanced representation of age and gender distributions in the dataset, prioritizing age distribution, which has been shown to affect model performance (ref).}
+        \caption{\label{tab:cobre_clinical_demographic}The subjects were selected to give a balanced representation of age and gender distributions in the dataset, prioritizing age distribution, which has been shown to affect model performance ~\cite{Oh2020}.}
         \begin{tabular*}{500pt}{@{\extracolsep\fill}lcccccc@{\extracolsep\fill}}
@@
 
-\subsubsection{intensity normalization}
+\subsubsection{Intensity normalization}
 

### 2025-02-15 src/utils/plot_svm_metrics.py

@@ -7,7 +7,8 @@ from sklearn.metrics import confusion_matrix, roc_curve, auc
 
 
 class SVMVisualizer:
-    def __init__(self, results_json_path, save_figs_path:str = "paper/figs"):
+
+    def __init__(self, results_json_path, save_figs_path: str = "paper/figs"):
         """
         Initializes the visualization class and loads classification results.
 
@@ -21,7 +22,8 @@ class SVMVisualizer:
     def _load_results(self):
         """Loads classification results from the JSON file."""
         if not os.path.exists(self.results_json_path):
-            raise FileNotFoundError(f"Results file not found: {self.results_json_path}")
+            raise FileNotFoundError(
+                f"Results file not found: {self.results_json_path}")
 
         with open(self.results_json_path, "r") as f:
             return json.load(f)
@@ -36,33 +38,68 @@ class SVMVisualizer:
         unique_classes = sorted(set(y_test))
 
         fig = plt.figure(figsize=(6, 5))
-        sns.heatmap(conf_matrix, annot=True, fmt='d', cmap='Blues', 
-                    xticklabels=unique_classes, yticklabels=unique_classes)
+        sns.heatmap(conf_matrix,
+                    annot=True,
+                    fmt='d',
+                    cmap='Blues',
+                    xticklabels=unique_classes,
+                    yticklabels=unique_classes)
         plt.xlabel('Predicted Labels')
         plt.ylabel('True Labels')
         plt.title('Confusion Matrix')
-        fig.savefig(self.save_figs_path+"cm"+".png")
-        #plt.show()
+        fig.savefig(self.save_figs_path + "cm" + ".png")
 
     def plot_roc_curve(self):
         """Plots the ROC curve."""
         y_test = np.array(self.results["y_test"])
-        test_scores = np.array(self.results["test_scores"])  # Scores should be probabilities
-        
+        test_scores = np.array(
+            self.results["test_scores"])  # Scores should be probabilities
+
         # Handle binary classification vs. multi-class
         if test_scores.ndim == 1:
             fpr, tpr, _ = roc_curve(y_test, test_scores)
         else:
-            fpr, tpr, _ = roc_curve(y_test, test_scores[:, 1])  # Only take positive class scores
+            fpr, tpr, _ = roc_curve(
+                y_test, test_scores[:, 1])  # Only take positive class scores
 
         roc_auc = auc(fpr, tpr)
 
[…]

### 2025-02-15 src/main.py
resolve conflict in main

@@ -94,8 +94,7 @@ def main():
         eda.save_raw_images_metrics()
         logger.info("EDA completed.")
     else:
-        logger.info("EDA was skipped.")    
-
+        logger.info("EDA was skipped.")
 
     # Step 1: Select the data (train, test)
     """
@@ -138,27 +137,28 @@ def main():
     Saves preprocessed test images in data/raw_nii/test_set/preprocessed.
     """
     if IS_PERFORM_PREPROCESSING:
-        preprocess_images(raw_nii_dir=RAW_NII_DATA_DIR,
-                        train_set_dir=TRAIN_SET_DIR,
-                        test_set_dir=TEST_SET_DIR,
-                        is_normalize=IS_NORMALIZE_WHEN_PREPROCESS,
-                        norm_metod=NORMALIZATION_METHOD,
-                        min_max_min_val=MIN_VAL,
-                        min_max_max_val=MAX_VAL,
-                        is_brain_extraction=IS_BRAIN_EXTRACTION,
-                        brain_extraction_modality=MODALITY,
-                        brain_extraction_verbose=VERBOSE,
-                        is_crop=IS_CROP,
-                        is_smooth=IS_SMOOTH,
-                        smooth_sigma=SIGMA,
-                        smooth_order=ORDER,
-                        smooth_mode=MODE,
-                        smooth_cval=CVAL,
-                        smooth_truncate=TRUNCATE,
-                        is_re_normalize_after_smooth=IS_RE_NORMALIZE_AFTER_SMOOTH,
-                        is_preprocess_test_set=IS_PREPROCESS_TEST_SET,
-                        output_dir=PREPROCESSED_DATA_DIR,
-                        )
+        preprocess_images(
+            raw_nii_dir=RAW_NII_DATA_DIR,
+            train_set_dir=TRAIN_SET_DIR,
+            test_set_dir=TEST_SET_DIR,
+            is_normalize=IS_NORMALIZE_WHEN_PREPROCESS,
+            norm_metod=NORMALIZATION_METHOD,
+            min_max_min_val=MIN_VAL,
+            min_max_max_val=MAX_VAL,
+            is_brain_extraction=IS_BRAIN_EXTRACTION,
+            brain_extraction_modality=MODALITY,
+            brain_extraction_verbose=VERBOSE,
+            is_crop=IS_CROP,
+            is_smooth=IS_SMOOTH,
+            smooth_sigma=SIGMA,
+            smooth_order=ORDER,
[…]

### 2025-02-15 src/main.py
pipeline tested

@@ -4,7 +4,7 @@ from utils.eda import SchizophreniaEDA
 from utils.dataset_preparation import prepare_dataset
 from utils.preprocessing import preprocess_images
 from utils.augment_images import augment_images
-from utils.feature_extractor import feature_extraction_pipeline
+#from utils.feature_extractor import feature_extraction_pipeline
 from utils.classifier import train_and_evaluate
 from utils.plot_svm_metrics import SVMVisualizer
 
@@ -13,7 +13,7 @@ def main():
     logger.info("The program started...")
 
     # Load configuration from a config file
-    config_data = OmegaConf.load('config/config.yaml')
+    config_data = OmegaConf.load('src/config/config.yaml')
 
     # Accessing parameters
     RAW_PT_DATA_DIR = config_data.base.raw_pt_data_dir
@@ -25,8 +25,8 @@ def main():
     IS_PERFORM_DATASET_PREPARATION = config_data.dataset_preparation.is_perform_dataset_preparation
     TRAIN_SET_DIR = config_data.dataset_preparation.train_set_dir
     TEST_SET_DIR = config_data.dataset_preparation.test_set_dir
-    TRAIN_SET_RATIO = config_data.dataset_preparation.train_set_ratio  # 0.24: 50 - train, 12 - test
-    IS_NORMALIZE_WHEN_PREPARING = config_data.dataset_preparation.is_normalize_when_preparing  # True False
+    TRAIN_SET_RATIO = config_data.dataset_preparation.train_set_ratio  # 0.80: 50 - train, 12 - test
+    IS_NORMALIZE_WHEN_PREPARING = config_data.dataset_preparation.is_normalize_when_preparing
     NORMALIZATION_METHOD = config_data.dataset_preparation.normalization_method  # min-max
 
     IS_PERFORM_PREPROCESSING = config_data.preprocessing_params.is_perform_preprocessing
@@ -38,7 +38,7 @@ def main():
     IS_RE_NORMALIZE_AFTER_SMOOTH = config_data.preprocessing_params.is_re_normalize_after_smooth
     IS_PREPROCESS_TEST_SET = config_data.preprocessing_params.is_preprocess_test_set
 
-    VOXEL_SIZE = config_data.preprocessing_params.voxel_size
+    VOXEL_SIZE = tuple(config_data.preprocessing_params.voxel_size)
     MIN_VAL = config_data.preprocessing_params.min_val
     MAX_VAL = config_data.preprocessing_params.max_val
     MODALITY = config_data.preprocessing_params.modality
[…]

### 2025-02-15 src/utils/augment_images.py
updated paths

@@ -2,7 +2,7 @@ import os
 import torch
 import random
 from logger import logger
-from src.utils.augmentations import apply_translation, apply_rotation, apply_gaussian_noise
+from utils.augmentations import apply_translation, apply_rotation, apply_gaussian_noise
 
 def augment_images(
         augmentations: list[tuple],

### 2025-02-15 src/utils/classifier.py

@@ -2,7 +2,7 @@ import os
 import json
 import numpy as np
 from logger import logger
-from src.models.svm import SVMClassifier
+from models.svm import SVMClassifier
 from sklearn.metrics import accuracy_score, classification_report, roc_auc_score, confusion_matrix
 
 

### 2025-02-15 src/utils/dataset_preparation.py

@@ -1,13 +1,11 @@
 import os
-import shutil
 import random
 import torch
 from logger import logger
-from src.utils.preprocess import load_nii, resample_image, normalize_data
+from utils.preprocess import load_nii, resample_image, normalize_data
 
 
 def prepare_dataset(train_ratio: float = 0.76,
-                    raw_pt_dir: str = "data/raw_pt",
                     raw_nii_dir: str = "data/raw_nii",
                     train_set_output_dir: str = "train_set",
                     test_set_output_dir: str = "test_set",
@@ -17,49 +15,36 @@ def prepare_dataset(train_ratio: float = 0.76,
                     min_max_min_val: float = 0,
                     min_max_max_val: float = 1) -> None:
     """
-    Splits MRI images into training and test sets while maintaining the same split across .pt and .nii formats.
+    Splits MRI images into training and test sets for .nii format.
     
-    The .pt files are copied into respective train/test folders for first training experiment, while
     .nii files are preprocessed (resampled, optionally normalized) before being saved in .pt format.
 
     Parameters:
         train_ratio (float): Proportion of data to use for training.
-        raw_pt_dir (str): Directory containing raw .pt images.
         raw_nii_dir (str): Directory containing raw .nii images.
         normalize (bool): Whether to normalize the .nii images.
         norm_method (str): Normalization method ("min-max" or other supported methods).
     """
     # Define train/test directories
-    train_pt_dir, test_pt_dir = os.path.join(raw_pt_dir, train_set_output_dir), os.path.join(raw_pt_dir, test_set_output_dir)
     train_nii_dir, test_nii_dir = os.path.join(raw_nii_dir, train_set_output_dir), os.path.join(raw_nii_dir, test_set_output_dir)
 
     # Create required directories
-    for directory in [train_pt_dir, test_pt_dir, train_nii_dir, test_nii_dir]:
+    for directory in [train_nii_dir, test_nii_dir]:
         os.makedirs(directory, exist_ok=True)
 
-    # Get list of .pt files and shuffle
-    pt_images = [f for f in os.listdir(raw_pt_dir) if f.endswith(".pt")]
-    random.shuffle(pt_images)
+    # Get list of .nii files and shuffle
+    nii_images = [f for f in os.listdir(raw_nii_dir) if f.endswith(".nii.gz")]
+    random.shuffle(nii_images)
 
     # Split dataset
-    split_idx = int(len(pt_images) * train_ratio)
[…]

### 2025-02-15 src/utils/feature_extractor.py

@@ -6,8 +6,8 @@ import pandas as pd
 from logger import logger
 import tqdm
 from torch.utils.data import DataLoader
-from src.utils.data_loader import MRIDataset
-from src.models.cnn import FeatureExtractor
+from utils.data_loader import MRIDataset
+from models.cnn import FeatureExtractor
 
 
 # Function to get image paths

### 2025-02-15 src/utils/plot_svm_metrics.py

@@ -2,7 +2,6 @@ import os
 import json
 import numpy as np
 import matplotlib.pyplot as plt
-import seaborn as sns
 from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay, roc_curve, auc
 
 

### 2025-02-15 src/utils/preprocess_validation.py

@@ -6,8 +6,7 @@ from logger import logger
 from skimage.filters import threshold_otsu
 from skimage.metrics import mean_squared_error
 from skimage.metrics import structural_similarity as ssim
-
-from src.utils.preprocess import get_data, match_dimensions
+from utils.preprocess import get_data, match_dimensions
 
 
 def plot_slices(data: Union[np.ndarray, nib.Nifti1Image],

### 2025-02-15 src/utils/preprocessing.py

@@ -1,7 +1,7 @@
 import os
 import torch
 from logger import logger
-from src.utils.preprocess import (
+from utils.preprocess import (
     normalize_data,
     extract_brain,
     crop_to_largest_bounding_box,

### 2025-02-15 src/utils/plot_svm_metrics.py
minor

@@ -1,5 +1,6 @@
 import os
 import json
+import seaborn as sns
 import numpy as np
 import matplotlib.pyplot as plt
 from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay, roc_curve, auc

### 2025-02-16 src/paper/paper.tex
updated paper

@@
 
-\textbf{Background}\\
+\tab\textbf{Background}\\
 Schizophrenia (SZ) is a chronic mental disorder that presents challenges in accurate diagnosis due to its complex nature.
@@
 \textbf{Methods}\\
-We utilized the COBRE dataset, which includes structural MRI data from 62 individuals, divided into 30 schizophrenia patients and 32 healthy controls.
-A support vector classifier (SVC) was employed to classify the data.
-Various preprocessing techniques—normalization, brain extraction, and cropping—and data augmentation methods—translation, rotation, and Gaussian noise addition—were applied to the MRI images.
+We utilized a subset of the COBRE dataset, which includes structural MRI data from 62 individuals, divided into 30 schizophrenia patients and 32 healthy controls.
+A support vector classifier (SVC) was employed to classify the images.
+Various preprocessing techniques --- resampling, normalization, brain extraction, cropping, smoothing --- and data augmentation methods -- translation, rotation, and Gaussian noise addition --- were applied to the MRI data.
 Model performance was assessed using accuracy, recall, precision, and F1-score.
@@
 \textbf{Results}\\
-Normalization and brain extraction improved classification accuracy, with cropping achieving the highest accuracy (0.57) and F1-score (0.69) among the preprocessing techniques.
+Normalization and brain extraction improved classification accuracy, with cropping achieving the highest accuracy (57\%) and F1-score (0.69) among the preprocessing techniques.
 Combining preprocessing methods did not always result in performance improvements and sometimes led to a decrease in accuracy.
+The combination of resampling, normalization, and brain extraction gave the best performance (accuracy of 55\% and F1-score of 0.69) among the combinations explored in this study.
 Data augmentation significantly enhanced model performance, with translation and rotation leading to the highest accuracy (0.79) and F1-score (0.80).
-However, Gaussian noise had a detrimental effect, reducing accuracy and F1-score.
+The data for the final model was preprocessed with the optimal combination, translated, and rotated at random (resampling + normalization + brain extraction + translation + rotation) and produced the highest recorded accuracy and recall scores (81\% and 0.98 respectively).
 
@@
 Our findings indicate that careful selection of preprocessing and augmentation techniques is critical for optimizing model performance in schizophrenia detection.
-While certain preprocessing steps like normalization and cropping are beneficial, combining them may not always yield additive improvements.
-Augmentation techniques like translation and rotation proved effective in boosting model performance, while Gaussian noise should be avoided.
+While certain preprocessing steps like cropping and smoothing are beneficial, combining them may not always yield additive improvements.
+Augmentation techniques like translation and rotation proved effective in boosting model performance, due to their ability to increase the data size.
 These results contribute to the development of more standardized preprocessing pipelines for medical imaging, which could enhance the reliability and reproducibility of machine learning models in clinical applications.
@@
 
-This project systematically evaluates preprocessing and augmentation techniques (e.g., smoothing, translation) to establish a robust pipeline for SZ detection from MRI.
+This study systematically evaluates preprocessing and augmentation techniques to establish a robust pipeline for SZ detection from MRI.
 Standardizing these practices could improve the reproducibility and clinical applicability of ML-based diagnostics.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{./figs/sample_volume.png} % test_pipeline.png
+    \includegraphics[width=0.8\textwidth]{./figs/sample_volume.png} %
     \caption{Three image slices representing the average of a 3D column in three views: dorsal, lateral, and superior respectively}\label{fig:raw_mri_volume_slices}
@@
 
-We experimented with different combinations of preprocessing and data augmentation techniques to determine which pipeline yields the best performance of the ML/DL models regarding recall rate, precision, and F1 score.
+We experimented with different preprocessing and data augmentation techniques, and their combinations, to determine which pipeline yields the best performance of the feature extraction model and a classifier regarding accuracy, recall rate, precision, and F1 score.
 
@@
 Table~\ref{tab:preprocessing_pipeline} presents the preprocessing techniques used and their respective hyperparameters and tools used.
-Figure~\ref{fig:test_pipeline} gives a comparison of a raw image and a preprocessed image, also indicating the image metrics.
+Figure~\ref{fig:slice_comparison} gives a comparison of a raw image and a preprocessed image, also indicating the image metrics.
 %Figure~\ref{fig:test_pipeline} gives a visual representation of the conducted tests.
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{./figs/slice_comparison.png} % test_pipeline.png
-    \caption{Panel (A) shows a slice of a raw scan.
-Panel (B) shows a slice of a preprocessed scan.
-For preprocessed scans, the standard deviation of SNR values is relatively low (\mbox{$SNR= 25 \pm 1$}), which ensures consistency in the image quality and prevents the model from making predictions based on possible artifacts.}\label{fig:test_pipeline}
+    \includegraphics[width=0.6\textwidth]{./figs/slice_comparison.png} % test_pipeline.png
+    \caption{Panel A shows a slice of a raw scan.
+Panel B shows a slice of a preprocessed scan.
+For preprocessed scans, the standard deviation of SNR values is relatively low (\mbox{$SNR= 25 \pm 1$}), which ensures consistency in the image quality and prevents the model from making predictions based on possible artifacts.}\label{fig:slice_comparison}
 \end{figure}
@@
 
-The first step of the pipeline is resampling the image to a standard resolution to ensure uniform input for the model and consistent noise and signal profiles.
+The first preprocessing technique employed was resampling the images to a standard resolution to ensure uniform input for the model and consistent noise and signal profiles.
 
@@
 
-The second step of the pipeline is normalization of voxel intensities to a consistent range to maintain the relative scale of the data and produce values in a fixed range to satisfy models expectations.
+The second technique normalization of voxel intensities to a consistent range to maintain the relative scale of the data and produce values in a fixed range to satisfy models expectations.
 The adopted method scales the data to a range of [0, 1].
@@
 
-Intensity distributions were inspected to ensure normalization has not distorted the data.
+Intensity distributions were inspected to ensure normalization did not distort the data.
 
@@
 
-The third step of the pipeline is a brain extraction. The purpose of this stage is to
+The third technique employed was a brain extraction. The purpose of this stage is to
 \begin{itemize}
-    \item remove non-brain regions in order to minimizes noise, making it easier for the network to focus on relevant features;
+    \item remove non-brain regions in order to minimize noise, making it easier for the network to focus on relevant features;
     \item subsequently enhance the network's performance, especially in a schizophrenia detection task where small differences in brain anatomy are crucial;
@@
 
-The fourth step of the pipeline is cropping the brain region to a consistent size across scans.
+The fourth technique was cropping the brain region to a consistent size across scans.
 
@@
 
-The fifth step of the pipeline is smoothing the brain region to reduce noise and variability introduced by acquisition artifacts, enhancing the overall quality of the images.
+The fifth technique was smoothing the brain region to reduce noise and variability introduced by acquisition artifacts, enhancing the overall quality of the images.
 This was done to ensure that the downstream tasks are more robust and the model training is more effective.
@@
 
-The sixth step of the pipeline is normalization of the smoothed image to address potential intensity changes and ensure that all images are on a comparable scale.
+The sixth technique was normalization of the smoothed image to address potential intensity changes and ensure that all images are on a comparable scale.
 
@@
         \centering
-        \caption{\label{tab:preprocessing_pipeline}Preprocessing techniques used in the study.}
+        \caption{\label{tab:preprocessing_pipeline}Preprocessing techniques and the corresponding parameters used in the study.}
         \begin{tabular*}{500pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}
@@
 
-The augmented samples were plotted to ensure the transformations don't distort anatomical features.
+The augmented samples were plotted to ensure the transformations did not distort anatomical features.
 
@@
 This input size corresponds to the size of the images in the ImageNet dataset that was used to train ResNet-18 and was selected for compatibility with the pretrained weights of the ResNet-18 model.
-Figure~\ref{fig:training_architecture} gives a representation of the training architecture and flow.
+Figure~\ref{fig:model_architecture} gives a representation of the training architecture and flow.
 %
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{./figs/model_architecture.png} % training_architecture.png
-    \caption{Training architecture and flow.}\label{fig:training_architecture}
+    \includegraphics[width=0.8\textwidth]{./figs/model_architecture.png} %
+    \caption{Overview of the training pipeline.
+Raw or preprocessed MRI slices are fed into a ResNet-18 model for feature extraction, where convolutional layers, batch normalization, and residual blocks process the input.
+The final feature vector, obtained after global average pooling and flattening, is passed to an SVM classifier with an RBF kernel for binary classification into schizophrenic and non-schizophrenic labels.}\label{fig:model_architecture}
 \end{figure}
@@
 The results presented in Table~\ref{tab:model_metrics_preprocessing} demonstrate the impact of various preprocessing methods on classification performance.
-The baseline model using raw images (0) achieved an accuracy of 0.51, with an F1-score of 0.63.
+The baseline model using raw images (0) achieved an accuracy of 51\%, with an F1-score of 0.63.
 Applying resampling (1) alone did not alter these results, suggesting that resampling alone does not introduce substantial benefits.
 
-Normalization (2) improved accuracy to 0.56 and F1-score to 0.67, indicating that standardizing intensity values enhances model performance.
-Brain extraction (3) yielded a slightly lower accuracy (0.54) but a notable increase in recall (0.78), which suggests that removing non-brain tissues helps capture more positive instances but may introduce a trade-off in precision.
+Normalization (2) improved accuracy to 56\% and F1-score to 0.67, indicating that standardizing intensity values enhances model performance.
+Brain extraction (3) yielded a slightly lower accuracy (54\%) but a notable increase in recall (0.78), which suggests that removing non-brain tissues helps capture more positive instances but may introduce a trade-off in precision.
 
-Cropping (4) led to the highest accuracy (0.57) and an F1-score of 0.69, highlighting its effectiveness in improving signal-to-noise ratio.
+Cropping (4) led to the highest accuracy (57\%) and an F1-score of 0.69, highlighting its effectiveness in improving signal-to-noise ratio.
 Smoothing (5) showed a moderate impact, improving results compared to the raw data but not outperforming other methods.
@@
 The combination of resampling, normalization, and brain extraction (1+2+3) resulted in the highest recall (0.80) while maintaining a strong F1-score (0.69).
-However, adding cropping (1+2+3+4) caused a drastic performance drop, with accuracy declining to 0.44 and F1-score to 0.54.
+However, adding cropping (1+2+3+4) caused a drastic performance drop, with accuracy declining to 44\% and F1-score to 0.54.
 The full pipeline (1+2+3+4+5) slightly improved over this but still underperformed compared to individual preprocessing steps.
@@
 		\centering
-		\caption{\label{tab:model_metrics_preprocessing}Desct}
+		\caption{\label{tab:model_metrics_preprocessing}Classification performance metrics for different preprocessing methods.
+Accuracy, recall, precision, and F1-score are reported for models trained on raw MRI slices and various preprocessing techniques.
+While normalization (2) and cropping (4) provided the highest accuracy (57\%), brain extraction (3) led to the highest recall (0.78).
+The combination of resampling, normalization, and brain extraction (1+2+3) resulted in the best recall (0.80) with a strong F1-score (0.69).
+However, adding cropping (1+2+3+4) and smoothing (1+2+3+4+5) led to performance degradation, suggesting that excessive preprocessing may reduce classification effectiveness.}
 		\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
 			\toprule
-		     \textbf{Preprocessing method} &\textbf{Accuracy} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
+		     \textbf{Preprocessing method} &\textbf{Accuracy (\%)} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
 			\midrule
-			Raw (0) & 0.51 & 0.67 & 0.60 & 0.63 \\
-			Resampling (1) & 0.51 & 0.67 & 0.60 & 0.63 \\
-			Normalization (2) & 0.56 & 0.69 & 0.64 & 0.67 \\
-			Brain extraction (3) & 0.54 & 0.78 & 0.60 & 0.68 \\
-			Cropping (4) & 0.57 & 0.76 & 0.63 & 0.69 \\
-			Smoothing (5) & 0.55 & 0.71 & 0.62 & 0.66 \\
-            1+2 & 0.57 & 0.69 & 0.64 & 0.67 \\
-            1+2+3 & 0.55 & 0.80 & 0.61 & 0.69 \\
-            1+2+3+4 & 0.44 & 0.53 & 0.56 & 0.54 \\
-            1+2+3+4+5 & 0.48 & 0.60 & 0.58 & 0.59 \\
+			Raw (0) & 51\% & 0.67 & 0.60 & 0.63 \\
+			Resampling (1) & 51\% & 0.67 & 0.60 & 0.63 \\
+			Normalization (2) & 56\% & 0.69 & 0.64 & 0.67 \\
+			Brain extraction (3) & 54\% & 0.78 & 0.60 & 0.68 \\
+			Cropping (4) & 57\% & 0.76 & 0.63 & 0.69 \\
+			Smoothing (5) & 55\% & 0.71 & 0.62 & 0.66 \\
+            1+2 & 57\% & 0.69 & 0.64 & 0.67 \\
+            1+2+3 & 55\% & 0.80 & 0.61 & 0.69 \\
+            1+2+3+4 & 44\% & 0.53 & 0.56 & 0.54 \\
+            1+2+3+4+5 & 48\% & 0.60 & 0.58 & 0.59 \\
 			\bottomrule
@@
 The baseline model using raw data (0) achieved an accuracy of 0.51 and an F1-score of 0.63.
-In contrast, applying translation (1) significantly improved both accuracy (0.75) and F1-score (0.79), with a notable increase in recall (0.84), indicating that translation is effective at boosting model performance and sensitivity.
+In contrast, applying translation (6) significantly improved both accuracy (75\%) and F1-score (0.79), with a notable increase in recall (0.84), indicating that translation is effective at boosting model performance and sensitivity.
 
-Rotation (2) further improved results, with the highest accuracy (0.79) and F1-score (0.80), alongside strong recall (0.81) and precision (0.78), demonstrating its ability to enhance both model robustness and overall performance.
+Rotation (7) further improved results, with the highest accuracy (79\%) and F1-score (0.80), alongside strong recall (0.81) and precision (0.78), demonstrating its ability to enhance both model robustness and overall performance.
 
-Gaussian noise (3), however, led to a substantial decrease in performance, with an accuracy of only 0.33 and a low F1-score of 0.40.
+Gaussian noise (8), however, led to a substantial decrease in performance, with an accuracy of only 33\% and a low F1-score of 0.40.
 This suggests that adding noise may negatively affect the model, potentially obscuring relevant patterns in the data.
-
 \begin{center}
@@
 		\centering
-		\caption{\label{tab:model_meetrics_augmentation}Desct}
+		\caption{\label{tab:model_meetrics_augmentation}Classification performance metrics for different data augmentation methods.
+Accuracy, recall, precision, and F1-score are reported for models trained on raw MRI slices and augmented datasets.
+Both translation (6) and rotation (7) significantly improved model performance, with rotation achieving the highest accuracy (79\%) and F1-score (0.80).
+In contrast, Gaussian noise (8) led to a severe drop in performance, reducing accuracy to 33\% and F1-score to 0.40.
+The final training setup, combining preprocessing with translation and rotation (1+2+3+6+7), achieved the best overall results, with 81\% accuracy and the highest recall (0.98), indicating the effectiveness of spatial transformations in improving classification.}
 		\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
 			\toprule
-			\textbf{Augmentation method} &\textbf{Accuracy} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
+			\textbf{Augmentation method} &\textbf{Accuracy (\%)} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
 			\midrule
 			Raw (0) & 0.51 & 0.67 & 0.60 & 0.63 \\
-			Translation (1) & 0.75 & 0.84 & 0.74 & 0.79 \\
-			Rotation (2) & 0.79 & 0.81 & 0.78 & 0.80 \\
-			Gaussian Noise (3) & 0.33 & 0.32 & 0.53 & 0.40 \\
+			Translation (6) & 75\% & 0.84 & 0.74 & 0.79 \\
+			Rotation (7) & 79\% & 0.81 & 0.78 & 0.80 \\
+			Gaussian Noise (8) & 33\% & 0.32 & 0.53 & 0.40 \\
+            Final training (1+2+3+6+7) & 81\% & 0.98 & 0.77 & 0.86 \\
 			\bottomrule
@@
 \end{center}
+
+The final training combined the optimal combination of preprocessing with translation and rotation techniques (1+2+3+6+7), surpassing the accuracy and recall of all the experiments (81\% and 0.98 respectively), with a precision value only surpassed by the rotation experiment.
+Figure~\ref{fig:eval_final_model} shows the output of the final training and Figure~\ref{fig:model_metrics_comparison} shows the comparison between the metrics of all the preprocessing and augmentation experiments.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{./figs/eval_final_model.png} % 
+    \caption{Preprocessing and augmentation steps: resampling + normalization + brain extraction + translation + rotation.
+Panel A: confusion matrix with recall of 0.98.
+Panel B: ROC curve with AUC of 0.86}\label{fig:eval_final_model}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{./figs/model_metrics_comparison.png} % 
+    \caption{Summary of the metrics for each preprocessing experiment.
+Panel A: comparison of accuracies across the experiments.
+The highest accuracy was recorded in the final experiment combining resampling, normalization, brain extraction, translation and rotation.
+Panel B: comparison of recall scores.
+The final experiment yielded the highest recall score.
+Panel C: comparison of precision scores.
+The rotation experiment produced the highest score, probably due to an increase in instances of the negative class after augmentation.
+The numbers along the y-axis represent preprocessing and augmentation scenarios: 0 --- raw images, 1 --- resampling, 2 --- normalization, 3 --- brain extraction, 4 --- cropping, 5 --- smoothing, 6 --- translation, 7 --- rotation, 8 --- Gaussian noise.}\label{fig:model_metrics_comparison}
+\end{figure}
 

### 2025-02-16 src/utils/eda.py
added eda data

@@ -11,8 +11,9 @@ class SchizophreniaEDA:
     """Performs exploratory data analysis (EDA) on the clinical data and calculates image metrics"""
 
     def __init__(self,
-                 path_to_clinical_data: str,
-                 path_to_raw_images: Optional[str] = None) -> None:
+                 path_to_clinical_data: str = "data/clinical_data.csv",
+                 path_to_raw_images: Optional[str] = "data/raw_pt",
+                 output_path: str = "data/eda") -> None:
         """
         Initializes the SchizophreniaEDA class with paths to clinical and raw imaging data.
 
@@ -27,6 +28,7 @@ class SchizophreniaEDA:
         self.schiz_df: pd.DataFrame = self._add_schiz_column()
         self._create_age_groups()
         self.path_to_raw_images: Optional[str] = path_to_raw_images
+        self.output_path = output_path
 
     def _add_schiz_column(self) -> pd.DataFrame:
         """
@@ -70,7 +72,7 @@ class SchizophreniaEDA:
         plt.legend(title='Diagnosis', bbox_to_anchor=[1, 0.9])
         plt.tight_layout()
         plt.show()
-        fig.get_figure().savefig(save_path)
+        fig.get_figure().savefig(self.output_path + "/" + save_path)
 
     def plot_gender_distribution(self, save_path='schiz_gen.png') -> None:
         """
@@ -94,7 +96,7 @@ class SchizophreniaEDA:
         plt.legend(title='Diagnosis', bbox_to_anchor=[1, 0.9])
         plt.tight_layout()
         plt.show()
-        fig.get_figure().savefig(save_path)
+        fig.get_figure().savefig(self.output_path + "/ " +save_path)
 
     def get_age_statistics(self) -> Dict[str, pd.Series]:
         """
@@ -110,6 +112,23 @@ class SchizophreniaEDA:
             "No Schizophrenia": df_ns["age"].describe()
         }
 
+    def save_age_statistics(self, filename: str = "age_statistics.csv"):
+        """
+        Computes age statistics and saves them to a CSV file.
+
+        Args:
+            filename (str): The name of the CSV file to save the statistics.
+        """
+        age_stats = self.get_age_statistics()
+
+        # Convert dictionary to DataFrame
+        df_stats = pd.DataFrame(age_stats)
+
+        # Transpose so that the statistics are rows and groups are columns
+        df_stats.to_csv(self.output_path + "/" + filename, index=True)
+
+        print(f"Statistics saved to {filename}")
+
[…]

### 2025-02-16 src/paper/paper.tex

@@
 
-\tab\textbf{Background}\\
+\textbf{Background}\\
 Schizophrenia (SZ) is a chronic mental disorder that presents challenges in accurate diagnosis due to its complex nature.
@@
 
-Among the individual preprocessing techniques, normalization resulted in the most substantial performance improvement, enhancing accuracy by 9.8\%.
+Among the individual preprocessing techniques, normalization resulted in the most substantial performance improvement, enhancing accuracy to 56\% (9.8\% increase compared to baseline).
 This can be attributed to the equalization of feature contributions, which allowed the model to make more accurate predictions.
@@
 However, combining preprocessing steps often led to diminished performance, especially when all techniques were applied together.
-The full preprocessing pipeline, including resampling, normalization, brain extraction, cropping, and smoothing, resulted in a performance drop, particularly in terms of accuracy and F1-score.
+While the best performance was achieved with resampling, normalization, and brain extraction, adding further steps, such as cropping and smoothing, resulted in a performance drop, particularly in terms of accuracy and F1-score.
 This suggests that certain preprocessing steps, when combined, may interfere with each other, creating distortions that prevent the model from learning the most pertinent features of the data.
@@
 Regarding data augmentation, both translation and rotation significantly enhanced model performance, with translation improving recall and rotation achieving the highest accuracy and F1-score.
-These results align with the literature, where such transformations are often used to increase the diversity of the training data, helping the model generalize better to unseen data.
+These results align with the literature, where such transformations are often used to increase the diversity of the training data, helping the model generalize better to unseen samples.
 Gaussian noise, however, severely degraded performance, suggesting that it may obscure important features in the MRI images, leading to poorer model predictions.
+
+The final training model, which combined the optimal preprocessing steps (resampling, normalization, and brain extraction) with translation and rotation (1+2+3+6+7), achieved the highest overall performance, with an accuracy of 81\%, recall of 0.98, and an F1-score of 0.86.
+This result demonstrates the effectiveness of combining well-selected preprocessing and augmentation techniques in enhancing classification performance.
+Notably, the final model's recall score suggests that it is highly sensitive to positive cases, which is crucial for medical applications where detecting true positives is a priority.
 
@@
 Overall, this study highlights the importance of standardizing preprocessing and augmentation techniques in medical imaging, particularly for tasks like schizophrenia detection.
-While certain methods demonstrated clear benefits, the combination of multiple preprocessing steps should be approached with caution, as it may lead to diminishing returns or reduced model performance.
-Further research is needed to explore the optimal combination of preprocessing techniques and to investigate the impact of more advanced augmentation methods on model accuracy and generalizability.
+The results suggest that the combination of multiple preprocessing steps should be approached with caution, as it may lead to diminishing returns or reduced model performance while a well-optimized combination of preprocessing and augmentation can significantly improve model performance.
+Further research is needed to explore additional augmentation strategies and evaluate their impact on model generalizability in larger and more diverse datasets.
 
@@
 
-In this study, we investigated the impact of various preprocessing and data augmentation techniques on the classification performance of schizophrenia from structural MRI images.
-Our results demonstrate that individual preprocessing methods, particularly normalization, brain extraction, and cropping, can positively influence model performance, improving accuracy, recall, and F1-score.
-However, combining these techniques did not always lead to enhanced results, suggesting that over-processing may introduce distortions that hinder model generalization.
+This study examined the impact of preprocessing and data augmentation on schizophrenia classification from structural MRI images.
+The best preprocessing setup (resampling + normalization + brain extraction) improved performance (55\% accuracy, 0.69 F1-score), but excessive preprocessing did not yield further gains.
 
-Data augmentation proved to be a valuable tool, with translation and rotation significantly improving model robustness and performance, particularly in recall and accuracy.
-Conversely, adding Gaussian noise resulted in a substantial performance decrease, indicating that it may obscure critical features in the data.
+Data augmentation had a stronger effect, with rotation achieving the highest accuracy (79\%) and F1-score (0.80).
+The final model, trained with optimized preprocessing and augmentation, achieved the best overall performance (81\% accuracy, 0.98 recall), highlighting the value of spatial transformations.
 
 The findings emphasize the importance of careful selection and application of preprocessing and augmentation methods in medical imaging tasks.
-While some techniques show clear benefits, the combination of multiple preprocessing steps must be handled cautiously to avoid diminishing returns.
-These results underscore the need for a standardized preprocessing pipeline to enhance the reliability, reproducibility, and clinical applicability of machine learning models for schizophrenia detection.
+The results underscore the need for a standardized preprocessing pipeline to enhance the reliability, reproducibility, and clinical applicability of machine learning models for schizophrenia detection.
 
-Further work is needed to refine preprocessing pipelines and explore more advanced data augmentation methods to improve the performance of ML-based schizophrenia diagnosis.
+Further work is needed to refine preprocessing pipelines and explore additional augmentation methods to improve the performance of ML-based schizophrenia diagnosis.
 Ultimately, this study contributes to the ongoing efforts to make ML-driven medical diagnostics more robust and applicable in real-world clinical settings.

### 2025-02-16 src/paper/paper.tex
minor changes

@@
 Panel B shows a slice of a preprocessed scan.
-For preprocessed scans, the standard deviation of SNR values is relatively low (\mbox{$SNR= 25 \pm 1$}), which ensures consistency in the image quality and prevents the model from making predictions based on possible artifacts.}\label{fig:slice_comparison}
+For preprocessed scans, the standard deviation of SNR values is relatively low (\mbox{$SNR= 25 \pm 2$}), which ensures consistency in the image quality and prevents the model from making predictions based on possible artifacts.}\label{fig:slice_comparison}
 \end{figure}
@@
 In contrast, applying translation (6) significantly improved both accuracy (75\%) and F1-score (0.79), with a notable increase in recall (0.84), indicating that translation is effective at boosting model performance and sensitivity.
-
-Rotation (7) further improved results, with the highest accuracy (79\%) and F1-score (0.80), alongside strong recall (0.81) and precision (0.78), demonstrating its ability to enhance both model robustness and overall performance.
-
-Gaussian noise (8), however, led to a substantial decrease in performance, with an accuracy of only 33\% and a low F1-score of 0.40.
-This suggests that adding noise may negatively affect the model, potentially obscuring relevant patterns in the data.
 \begin{center}
@@
 \end{center}
+Rotation (7) further improved results, with the highest accuracy (79\%) and F1-score (0.80), alongside strong recall (0.81) and precision (0.78), demonstrating its ability to enhance both model robustness and overall performance.
+
+Gaussian noise (8), however, led to a substantial decrease in performance, with an accuracy of only 33\% and a low F1-score of 0.40.
+This suggests that adding noise may negatively affect the model, potentially obscuring relevant patterns in the data.
+\begin{figure}
+    \centering
+    \includegraphics[width=1.0\textwidth]{./figs/eval_final_model.png} % 
+    \caption{Model evaluation for the final training.
+Preprocessing and augmentation steps: resampling + normalization + brain extraction + translation + rotation.
+Panel A: confusion matrix with recall of 0.98.
+Panel B: ROC curve with AUC of 0.86}\label{fig:eval_final_model}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{./figs/eval_final_model.png} % 
-    \caption{Preprocessing and augmentation steps: resampling + normalization + brain extraction + translation + rotation.
-Panel A: confusion matrix with recall of 0.98.
-Panel B: ROC curve with AUC of 0.86}\label{fig:eval_final_model}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{./figs/model_metrics_comparison.png} % 
-    \caption{Summary of the metrics for each preprocessing experiment.
+    \includegraphics[width=1.0\textwidth]{./figs/model_metrics_comparison.png} % 
+    \caption{Summary of the metrics across different experiments.
 Panel A: comparison of accuracies across the experiments.

### 2025-02-17 src/paper/paper.tex

@@
 The dataset was split into a training set (80\%) and a validation set (20\%).
+Preprocessing, including resampling, normalization, and brain extraction, was applied before training.
+Data augmentation (translation and rotation) was performed on the training set to improve model generalization, while the validation set remained unaltered.
 
-Accuracy, recall, precision, and F1-score were calculated to asses the performance of the SVC.
+A Support Vector Classifier (SVC) was trained with an RBF kernel, kernel coefficient \mbox{$gamma=0.0001$} and optimized using regularization parameter \mbox{$C=100$}.
+Accuracy, recall, precision, and F1-score were used to assess performance of the Classifier.
 

### 2025-02-18 src/paper/paper.tex
update paper.tex

@@
 
-The data obtained originates from the Center of Biomedical Research Excellence (COBRE) dataset in NifTI format. It includes data from 62 individuals, 30 with schizophrenia and 32 healthy controls.
+The data obtained originates from the Center of Biomedical Research Excellence (COBRE) dataset in NifTI format.
+It includes data from 62 individuals, 30 with schizophrenia and 32 healthy controls.
+The subset used consists of 62 three-dimensional (3D) MRI scans with 192 slices each, bringing the total number of slices to 11,904.
+The list of subjects whose images were used can be obtained from the \href{https://github.com/Eager1Beaver/schizo/blob/main/data/clinical_data.csv}{clinical data}
 
 The data from SchizConnect was filtered based on MRI field strength and clinical diagnosis.
-We retrieved only images captured with 3T MRI scans.
+We retrieved only T1-weighted images (MPRAGE sequence) acquired using 3T MRI scanners.
 This was done to ensure uniformity of the data obtained.

### 2025-02-18 src/paper/paper.tex
typos fixed

@@
     \includegraphics[width=0.8\textwidth]{./figs/sample_volume.png} %
-    \caption{Three image slices representing the average of a 3D column in three views: dorsal, lateral, and superior respectively}\label{fig:raw_mri_volume_slices}
+    \caption{Three image slices representing the average of a 3D column in three views: dorsal, lateral, and superior respectively.}\label{fig:raw_mri_volume_slices}
 \end{figure}
